---
title: "01c_allocate_unknown_trade_fish_consumption"
output: html_document
date: "2023-05-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objective: The fish trade matrix for human consumption has a number of data points that are "unknown" in the source country, habitat, method, and nceas_group columns. Unknowns arise from the error term or cases where products move through more than 2 intermediate countries. We will proportionately bump up all marine flows with the marine-unknown volume, all inland flows with the inland-unknown, and then all aquatic food flows with the unknown-unknown. This will keep the total consumption constant without having to assume a global average for the pressure values when integrated.

## Setup
```{r}
library(tidyverse)
```

### File paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
matrix_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/trade/aquatic_food_trade/"
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01c_allocate_unknown_trade_fish_consumption/"
```

### Read in data
```{r}
seafood_matrix <- read.csv(paste0(matrix_path, "20230420_nceas_consumption.csv"))
```

### Identify combinations of unknown data
```{r}
# Identify typologies of unknown 
unknown_types <- seafood_matrix %>%
  filter(source_country_iso3c == "unknown"|       
           nceas_group == "unknown"|
           habitat == "unknown"|
           method == "unknown") %>%
  select(source_country_iso3c, nceas_group, habitat, method) %>%
  mutate(
    source_country_iso3c = case_when(source_country_iso3c != "unknown" ~ "known",
                                     TRUE ~ "unknown"),
    nceas_group = case_when(nceas_group != "unknown" ~ "known",
                            TRUE ~ "unknown"),
    habitat = case_when(habitat != "unknown" ~ "known",
                        TRUE ~ "unknown"),
    method = case_when(method != "unknown" ~ "known",
                       TRUE ~ "unknown")
    ) %>%
  group_by(source_country_iso3c, nceas_group, habitat, method) %>%
  tally()

# Add typology group to source-consumption data frame
seafood_matrix <- seafood_matrix %>%
  mutate(unknown_type = case_when(
    (source_country_iso3c != "unknown" &
       nceas_group == "unknown" &
       habitat != "unknown" &
       method != "unknown") ~ "species_unknown",
    (source_country_iso3c == "unknown" &
     nceas_group == "unknown" &
     habitat != "unknown" &
     method == "unknown") ~ "habitat_only",
    (source_country_iso3c == "unknown" &
     nceas_group == "unknown" &
     habitat == "unknown" &
     method == "unknown") ~ "all_unknown",
    TRUE ~ "known"
  ))
```


### Unknown type 1: nothing known about source, habitat, method, or species group
```{r}
# Use proportions of all known consumption
all_unknown_prop_df <- seafood_matrix %>%
  # Remove rows where any information is unknown
  filter(source_country_iso3c != "unknown",       
         nceas_group != "unknown",
         habitat != "unknown",
         method != "unknown") %>%
  # Get proportions for each consuming country
  group_by(consumer_iso3c) %>%
  mutate(all_unknown_prop = consumption_live_weight_t/sum(consumption_live_weight_t)) 

all_unknown <- all_unknown_prop_df %>% 
  left_join(seafood_matrix %>%
              # Filter to rows where all information is unknown
              filter(unknown_type == "all_unknown") %>%
              select(consumer_iso3c, "all_unknown_t" = "consumption_live_weight_t"),
            by = "consumer_iso3c") %>%
  mutate(consumption_live_weight_t = all_unknown_prop*all_unknown_t) %>%
  select(-all_unknown_prop, -all_unknown_t)
```

# Unknown type 2: only habitat known
```{r}
# Use proportions of all known consumption by habitat
habitat_known_prop_df <- seafood_matrix %>%
  # Remove rows where habitat is unknown
  filter(habitat != "unknown") %>%
  # Get proportions for known habitats for each consuming country
  group_by(consumer_iso3c, habitat) %>%
  mutate(habitat_known_prop = consumption_live_weight_t/sum(consumption_live_weight_t)) 

habitat_known <- habitat_known_prop_df %>% 
  left_join(seafood_matrix %>%
              # Filter to rows where all information is unknown other than habitat
              filter(unknown_type == "habitat_only") %>%
              select(consumer_iso3c, habitat, "habitat_known_t" = "consumption_live_weight_t"),
            by = c("consumer_iso3c", "habitat")) %>%
  mutate(consumption_live_weight_t = habitat_known_prop*habitat_known_t) %>%
  select(-habitat_known_prop, -habitat_known_t)
```

# Unknown type 3: only species group is unknown
```{r}
# Use proportions of all known consumption by source, habitat, and method
species_unknown_prop_df <- seafood_matrix %>%
  # Remove rows where nceas group  is unknown
  filter(nceas_group != "unknown") %>%
  # Get proportions each consuming country by source, habitat, and method
  group_by(consumer_iso3c, source_country_iso3c, habitat, method) %>%
  mutate(species_unknown_prop = consumption_live_weight_t/sum(consumption_live_weight_t)) 

species_unknown <- seafood_matrix %>% 
  # Filter to rows where only species information is unknown
  filter(unknown_type == "species_unknown") %>%
  select(consumer_iso3c, source_country_iso3c, habitat, method,
         "species_unknown_t" = "consumption_live_weight_t") %>%
  left_join(species_unknown_prop_df,
            by = c("consumer_iso3c", "source_country_iso3c", "habitat", "method")) %>%
  mutate(consumption_live_weight_t = species_unknown_prop*species_unknown_t) %>%
  select(-species_unknown_prop, -species_unknown_t)
```

### Combine all data frames with the known instances
```{r}
# Join all together
seafood_matrix_new <- seafood_matrix %>%
  filter(unknown_type == "known") %>%
  bind_rows(all_unknown) %>%
  bind_rows(habitat_known) %>%
  bind_rows(species_unknown)
```

### Check consumption
```{r}
seafood_check <- seafood_matrix %>%
  group_by(consumer_iso3c) %>%
  summarise(consumption_original = sum(consumption_live_weight_t, na.rm = TRUE)) %>%
  left_join(seafood_matrix_new %>%
              group_by(consumer_iso3c) %>%
              summarise(consumption_new = sum(consumption_live_weight_t, na.rm = TRUE)),
            by = "consumer_iso3c") %>%
  mutate(diff = consumption_new-consumption_original) %>%
  mutate(perc_diff = 100*(diff/consumption_original)) %>%
  arrange(desc(abs(perc_diff)))
```

Note: The original consumption does not match the disaggregated consumption 1:1 due to cases where no proportion can be calculated. For example, if all marine capture sourced from a given country is unknown, a proportion cannot be calculated for known species groups. We could consider using the proportions of the source country to estimate this. Currently, all but 5 countries lose less than 1% of consumption. 

### Final cleaning
```{r}
seafood_matrix_new <- seafood_matrix_new %>%
  # Remove small values created by proportions
  filter(consumption_live_weight_t > 10^-9) %>%
  # Remove fmfo as this is handled separately
  filter(nceas_group != "fomf") %>%
  # Select specific columns and save
  select(consumer_iso3c, source_country_iso3c, habitat, method, nceas_group, consumption_live_weight_t)

write_csv(seafood_matrix_new, paste0(save_path, "fish_trade_matrix_w_unknowns_allocated.csv"))

```

