---
title: "connect_2017_fw_aquaculture_to_pressures"
author: "Haley Epperly"
date: "2022-11-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```

File paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
aquatic_trade_data_path <- paste0(raw_data_path,"trade/aquatic_food_trade/")
```

## Method update with new data
```{r}
trade_matrix_raw <- read_csv(paste0(aquatic_trade_data_path, "20230420_nceas_consumption.csv")) # read in the updated data

fw_trade_clean <- trade_matrix_raw %>% 
  filter(!source_country_iso3c == "unknown",              # filter out data with unknown source country
         habitat == "inland" & method == "aquaculture")   # only want inland aquaculture to accurately allocate freshwater (inland) aquaculture pressures
```

For each fish producing/catching country, calculate the proportion of each organism (using nceas categories) that they trade with each consuming country (e.g., trade 5% of tuna with country B, 7% with country C, etc.)
```{r}
fw_trade_proportions <- fw_trade_clean %>% 
  group_by(source_country_iso3c, consumer_iso3c, nceas_group) %>% 
  summarise(traded_tons = sum(consumption_live_weight_t, na.rm = TRUE)) %>%     # sum the total tons of nceas_group for each consumer country from the source
  ungroup() %>%                                                                 # ungroup for further calcs
  group_by(source_country_iso3c, nceas_group) %>% 
  mutate(sum_caught = sum(traded_tons, na.rm = TRUE)) %>%                       # sum the total tons farmed by species for each source country
  ungroup() %>%                                                                 # ungroup for further calcs
  mutate(proportion_traded = traded_tons/sum_caught)                            # calculate proportions of catch from source country goes to consumer country

# check that proportions sum to 1 for each animal product in each producing country
# check_props <- fw_trade_totals %>% 
#   group_by(source_country_iso3c, nceas_group) %>% 
#   summarise(sum = sum(proportion_traded))
# 
# unique(check_props$sum)                                                         # only 1s so the proportions are calculated correctly
```



## Haley Method with old data below
Objective: Use the 2017 seafood trade matrix created by Jessica and the inland aquaculture pressures calculated by Mel to calculate the environmental pressures from traded freshwater aquaculture products. We should have associated pressures for each product between each pair of trading partners.

Read in fisheries trade matrix from Jessica
```{r}
tm <- read_csv("/home/shares/food-systems/social_justice_projects/_raw_data/aquatic_food_trade/20221011_nceas_ARTIS_snet.csv")

# subset to 2017 and inland aquaculture and remove rows with "export error" in dom source
tm_clean <- tm %>% 
  filter(year == 2017) %>% 
  filter(c(habitat == "inland" & method == "aquaculture")) %>% 
  filter(!dom_source == "error export")

# check data
table(tm_clean$dom_source)
table(tm_clean$year)

# remove jellyfish - Gage says jellfish were not included in pressures or mariculture or fisheries data
tm_clean <- tm_clean %>% 
  filter(!nceas_group == "jellyfish")
```

For each fish producing/catching country, calculate the proportion of each organism (using nceas categories) that they trade with each consuming country (e.g., trade 5% of tuna with country B, 7% with country C, etc.)
```{r}
# remove intermediate countries in trade matrix
tm_o_d <- tm_clean %>% 
  select(c(1, 3:4, 9)) %>% 
  group_by(source_country_iso3c, importer_iso3c, nceas_group) %>% 
  mutate(trade_weight = sum(live_weight_t)) %>% 
  select(-live_weight_t) %>% 
  unique() %>% 
  ungroup()

# check data
sum(tm_clean$live_weight_t)
sum(tm_o_d$trade_weight)

# look at one trading pair example
test3 <- tm_o_d %>% 
  filter(source_country_iso3c == "CHN" & importer_iso3c == "CHL")

# sum total amount of each organism a country produces/catches
fish_trade_sum <- tm_o_d %>% 
  group_by(source_country_iso3c, nceas_group) %>% 
  mutate(sum_caught = sum(trade_weight))

# calculate proportion they trade to each country (including consuming themselves)
fish_trade_prop <- fish_trade_sum %>% 
  mutate(prop_traded = trade_weight/sum_caught)

# check that proportions sum to 1 for each animal product in each producing country
check_props <- fish_trade_prop %>% 
  group_by(source_country_iso3c, nceas_group) %>% 
  summarise(sum = sum(prop_traded))

# all sum to 1 so that looks good
range(check_props$sum, na.rm=TRUE)
length(which(is.na(check_props$sum))) # 0 NAs
```

## Below the new data is incorporated to the workflow

Read in freshwater aquaculture pressures data from Mel
https://github.com/jagephart/food_trade_ej/blob/main/scripts/fw_aquaculture_pressures/output/fw_aquaculture_pressures.csv
```{r}
pressure <- read_csv("https://raw.githubusercontent.com/jagephart/food_trade_ej/main/scripts/fw_aquaculture_pressures/output/fw_aquaculture_pressures.csv?token=GHSAT0AAAAAABXM6CBM4SXKH6YX3TTMUDRYY3WPXTQ")
```


Compare species groups - these are extremely different. 
```{r}
table(pressure$stressor_group)
table(fish_trade_prop$nceas_group)
```


Investigating salmon to make sure Salmons, trouts, smelts are all inland aquaculture and don't include marine species.
Salmonids are all marine aquaculture, so Salmons, trouts, smelts is probably just inland.
```{r}
salmonid <- tm %>% 
  filter(nceas_group=="salmonids")
salmon <- tm %>% 
  filter(nceas_group=="Salmons, trouts, smelts")
```


From Mel:
Carps, barbels and other cyprinids = oth_carp + hypoph_carp
Miscellaneous diadromous fishes = misc_diad + milkfish
Salmons, trouts, smelts = trout
Miscellaneous freshwater fishes = catfish
Tilapias and other cichlids = tilapia

Re-label pressures data to categories used in trade data (pressures data is more disaggregated)
```{r}
pressure_relabel <- pressure %>% 
  mutate(nceas_group = case_when(stressor_group == "catfish" ~ "Miscellaneous freshwater fishes",
                                 stressor_group == "hypoph_carp" ~ "Carps, barbels and other cyprinids",
                                 stressor_group == "milkfish" ~ "Miscellaneous diadromous fishes",
                                 stressor_group == "misc_diad" ~ "Miscellaneous diadromous fishes",
                                 stressor_group == "oth_carp" ~ "Carps, barbels and other cyprinids",
                                 stressor_group == "tilapia" ~ "Tilapias and other cichlids",
                                 stressor_group == "trout" ~ "Salmons, trouts, smelts",
                                 TRUE ~ as.character(NA)))
summary(pressure_relabel)

pressure_relabel_2 <- pressure_relabel %>% 
  mutate(pressure = case_when(pressure == "GHG" ~ "ghg",
                              pressure == "nutrients" ~ "nutrient",
                              TRUE ~ pressure))
summary(pressure_relabel_2)
```


Sum pressures for the same nceas category
```{r}
pressure_agg <- pressure_relabel_2 %>% 
  group_by(iso3c, country_name, pressure, nceas_group) %>% 
  summarise(sum_pressure_value = sum(stressor_value))

# check data
sum(pressure_agg$sum_pressure_value)
sum(pressure_relabel_2$stressor_value)
```


Subset trade to relevant categories for which we have pressures data (removes ~ 25% of rows)
```{r}
fish_trade_sub <- fish_trade_prop %>% 
  filter(nceas_group == "Miscellaneous freshwater fishes" | nceas_group == "Carps, barbels and other cyprinids" |
           nceas_group == "Miscellaneous diadromous fishes" | nceas_group == "Tilapias and other cichlids" |
           nceas_group == "Salmons, trouts, smelts")
```


To get to cumulative pressure, divide each pressure by the global pressure (to standardize) and then sum the four pressures for each animal product
https://github.com/OHI-Science/food_systems/blob/master/_analysis/rescale_values.csv
- go to this github link, click "Raw", then copy url and paste below to read in csv. 
This link needs to be updated each time we read it in to R!
```{r}
# read in data on total global pressures
total_global_pressure <- read_csv("https://raw.githubusercontent.com/OHI-Science/food_systems/master/_analysis/rescale_values.csv?token=GHSAT0AAAAAABXM6CBNMMDINGAOMTDKGFCKY3WR7GA")

# join the fish pressures with the total global pressures
sum_pressures <- left_join(pressure_agg, total_global_pressure, by = "pressure")

# divide each pressure by the global total pressure to standardize - this includes total pressures from all crops and animal products
sum_pressures_1 <- sum_pressures %>% 
  mutate(prop_global_pressure = sum_pressure_value/global_total)

# sum the four pressures for each crop
sum_pressures_and_cumu <- sum_pressures_1 %>% 
  group_by(iso3c, country_name, nceas_group) %>% 
  mutate(cumu_pressure = sum(prop_global_pressure)) %>% 
  ungroup()

sum(sum_pressures_and_cumu$cumu_pressure) # 0.3585281

# check: make sure cumulative pressures looks right (should have only one cumu pressure value for each crop product in each country)
test1 <- sum_pressures_and_cumu %>% 
  filter(country_name == "Brazil")
test2 <- sum_pressures_and_cumu %>% 
  filter(country_name == "Australia")

# change formatting so all in wide format (right now only cumulative pressure is in wide format and other pressures are in long format)
sum_pressures_and_cumu_wide <- sum_pressures_and_cumu %>% 
  select(!c(global_total, prop_global_pressure)) %>% 
  pivot_wider(names_from = pressure, values_from = sum_pressure_value)
```


Join the trade matrix with pressures df and look at missing values
```{r}
fish_trade_pressures_w_missing <- full_join(fish_trade_sub, sum_pressures_and_cumu_wide, by = c("source_country_iso3c" = "iso3c", "nceas_group" = "nceas_group"))

# check missing trade data (have pressures but no trade data)
missing_values <- fish_trade_pressures_w_missing %>% 
  filter(is.na(prop_traded))

# note: there are 23 rows of data that are included in sum_pressures_and_cumu_wide, but not in fish_trade_sub
# this means that we have 2017 pressures data for these items in these countries, but they are not in Jessica's trade matrix.

# for how many rows do we have trade data but not pressures data?
missing_pressures <- fish_trade_pressures_w_missing %>% 
  filter(is.na(cumu_pressure))

# note: there are 3,129 rows of data that are included in fish_trade_sub, but not in sum_pressures_and_cumu_wide
# this means that we have 2017 trade data for these items in these countries, but we don't have pressures data for them

# sum of pressures from missing values - we can use this as a check later
summary_missing_pressures <- missing_values %>% 
  summarise_at(vars(cumu_pressure:water), sum, na.rm = TRUE)

# left_join instead so we don't include those rows where we only have pressures data, but no trade
fish_trade_pressures <- left_join(fish_trade_sub, sum_pressures_and_cumu_wide, by = c("source_country_iso3c" = "iso3c", "nceas_group" = "nceas_group"))
```


Multiply proportion traded between each country with the total amount of pressures for that product produced in that country. Then you can get the value of pressures (4 individual plus cumulative) for each product trade between 2 countries.
```{r}
fish_trade_pressures_1 <- fish_trade_pressures %>% 
  mutate(trade_cumu_pressure = prop_traded*cumu_pressure) %>% 
  mutate(trade_ghg = prop_traded*ghg) %>% 
  mutate(trade_water = prop_traded*water) %>% 
  mutate(trade_nutrient = prop_traded*nutrient) %>% 
  mutate(trade_disturbance = prop_traded*disturbance)

# subset to remove extra variables
fish_trade_pressures_cut <- fish_trade_pressures_1 %>% 
  rename(Producer_country = country_name) %>% 
  select(-c(8:12))

# reorganize
fish_trade_pressures_cut <- fish_trade_pressures_cut[,c(1, 7, 2:6, 8:12)]

# save df
write_csv(fish_trade_pressures_cut, "/home/shares/food-systems/social_justice_projects/food_trade_ej/data/inland_aquaculture_products_trade_matrix_w_pressures.csv")

# check to see if it looks right
alb_test <- fish_trade_pressures_1 %>% 
  filter(source_country_iso3c == "ALB") %>% 
  filter(nceas_group == "Salmons, trouts, smelts")
sum(alb_test$trade_weight) # 796.3468 (matches sum_caught in df)
sum(alb_test$prop_traded) #1

# make sure total value of "Salmons, trouts, smelts" caught in ALB still matches original data
alb_check <- tm_o_d %>% 
  filter(source_country_iso3c == "ALB") %>% 
  filter(nceas_group == "Salmons, trouts, smelts")
sum(alb_check$trade_weight) # 796.3468 (matches ago_test)

# these values should match the rows labeled cumu_pressure, disturbance, ghg, nutrient, and water in alb_test
sum(alb_test$trade_cumu_pressure)
sum(alb_test$trade_disturbance)
sum(alb_test$trade_ghg)
sum(alb_test$trade_nutrient)
sum(alb_test$trade_water)
```



