---
title: "Freshwater aquaculture pressures"
output: html_document
date: "2022-11-02"
editor_options: 
  chunk_output_type: console
---

GOAL: Calculate pressures for freshwater aquaculture by country and taxa group.
```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(countrycode)
```

# FAO production data

Format these data to include only the years of interest
IN = freshwater
BW = brackishwater
MA = marine
Al = all environments
```{r}
raw_production <- read_csv("/home/shares/food-systems/social_justice_projects/_raw_data/fao_aquaculture/Aquaculture_Quantity.csv") %>%
  filter(ENVIRONMENT.ALPHA_2_CODE=="IN") %>%
  filter(PERIOD %in% 2017) %>%
  select(UN_Code = COUNTRY.UN_CODE, species_code = SPECIES.ALPHA_3_CODE, PERIOD, VALUE) %>%
  group_by(UN_Code, species_code) %>%
  summarize(tonnes_lw = mean(VALUE, na.rm=TRUE)) %>%
  ungroup()

country <- read_csv("/home/shares/food-systems/social_justice_projects/_raw_data/fao_aquaculture/CL_FI_COUNTRY_GROUPS.csv") %>%
  dplyr::select(UN_Code, UN_country_name = Name_En)

sum(duplicated(country$UN_Code))

species <- read_csv("/home/shares/food-systems/social_justice_projects/_raw_data/fao_aquaculture/CL_FI_SPECIES_GROUPS.csv") %>%
  dplyr::select(species_code = `3A_Code`, Taxonomic_Code, taxa_name = Name_En, Major_Group_En, ISSCAAP_Group_En, Scientific_Name) %>%
  unique
sum(duplicated(species$species_code))

setdiff(raw_production$species_code, species$species_code)
setdiff(raw_production$UN_Code, country$UN_Code)

production <- left_join(raw_production, country, by="UN_Code") %>%
  left_join(species, by="species_code") %>%
    mutate(UN_country_name = ifelse(UN_country_name=="Türkiye", "Turkey", UN_country_name)) %>%
  mutate(iso3c = countrycode(UN_country_name, origin="country.name", destination = "iso3c"),
                       country_name = countrycode(iso3c, origin= "iso3c", destination = "country.name")) %>%
  filter(!is.na(iso3c)) %>% # these are older country designations with no production for years of interest
  filter(tonnes_lw > 0)

```


# Assigning taxonomic categories

NOTE: Considered including the "Sturgeons, paddlefishes" as "Tilapias and other cichlids", but didn't due to Jessica's comments:
- I suspect a lot of that production is for caviar and the pressures for tilapia probably doesn't apply well.
- It also looks like there is quite a bit of production of juveniles for stocking recreational fisheries.
```{r}
groups <- read_csv(here("scripts/fw_aquaculture_pressures/inputs/taxa_isscaap_groupings.csv")) %>%
  filter(environment=="freshwater")
unique(groups$isscaap_group)

intersect(production$ISSCAAP_Group_En, groups$isscaap_group)
in_fao <- setdiff(production$ISSCAAP_Group_En, groups$isscaap_group)
setdiff(groups$isscaap_group, production$ISSCAAP_Group_En)


tmp <- filter(production, ISSCAAP_Group_En %in% in_fao)
unique(tmp$ISSCAAP_Group_En)

filter(production, ISSCAAP_Group_En %in% "Pearls, mother-of-pearl, shells" ) %>% data.frame()

## change some groupings here
production_update <- production %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Shrimps, prawns", "Freshwater crustaceans", ISSCAAP_Group_En)) %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Miscellaneous coastal fishes" & taxa_name =="Obscure pufferfish", "Miscellaneous freshwater fishes", ISSCAAP_Group_En)) %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Miscellaneous coastal fishes" & taxa_name =="Flathead grey mullet", "Miscellaneous freshwater fishes", ISSCAAP_Group_En)) %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Miscellaneous coastal fishes" & taxa_name =="Mullets nei", "Miscellaneous freshwater fishes", ISSCAAP_Group_En)) 

intersect(production_update$ISSCAAP_Group_En, groups$isscaap_group)
in_fao <- setdiff(production_update$ISSCAAP_Group_En, groups$isscaap_group)
setdiff(groups$isscaap_group, production$ISSCAAP_Group_En)


## now make some adjustments to match Jessica's groups

hypoph_carp <- unique(grep("Hypophthalmichthys", production_update$Scientific_Name, value=TRUE))

production_update <- production_update %>%
  mutate(stressor_group = ifelse(taxa_name %in% c("Bighead carp", "Silver carp", "Silver, bighead carps nei"), "hypoph_carp", NA)) %>%
  mutate(stressor_group = ifelse(Scientific_Name %in% hypoph_carp, "hypoph_carp", stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Carps, barbels and other cyprinids" & is.na(stressor_group), "oth_carp", stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Tilapias and other cichlids" & is.na(stressor_group), "tilapia", stressor_group)) %>%
 # mutate(stressor_group = ifelse(taxa_name %in% c("Asian clam", "Chinese pond mussel") & is.na(stressor_group), "bivalves", stressor_group))
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Miscellaneous freshwater fishes" & is.na(stressor_group), "catfish", stressor_group)) %>%  # most are catfish, see Jessica note in issue
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Salmons, trouts, smelts" & is.na(stressor_group), "trout", stressor_group)) %>%
  mutate(stressor_group = ifelse(taxa_name == "Milkfish" & is.na(stressor_group), "milkfish", stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Miscellaneous diadromous fishes" & is.na(stressor_group), "misc_diad", stressor_group))

tmp <- filter(production_update, ISSCAAP_Group_En %in% "Miscellaneous diadromous fishes") %>% data.frame()
tmp <- filter(production_update, is.na(stressor_group)) %>% 
  select(ISSCAAP_Group_En) %>% unique()

## accounts for about 91% of production
sum(production_update$tonnes_lw[!is.na(production_update$stressor_group)])/sum(production_update$tonnes_lw)

## summarize by Jessica's groupings
tonnes_lw <- production_update %>%
  filter(!is.na(stressor_group)) %>%
  group_by(iso3c, country_name, stressor_group) %>%
  summarize(tonnes_lw = sum(tonnes_lw)) %>%arrange(-tonnes_lw)

```

## Join with on-farm stressors
Units here are kg N/P per tonne, m3 water per tonne, kgC02eq per tonne, m2a. 
```{r}
pressures <- read.csv("/home/shares/food-systems/social_justice_projects/_raw_data/fresh_water_aqua_pressures_Jessica/SI_stressor_results_on_off_farm.csv") %>%
  filter(weight=="live") %>%
  filter(production=="aquaculture") %>%
  filter(source == "on-farm") %>%
  filter(allocation == "mass") %>%
  dplyr::select(stressor_group = taxa, stressor, median)

conversions <- data.frame(stressor = c("GHG", "N", "P", "Water", "Land"),
                          conversion = c(0.001, 0.001, 0.001, 1, 0.000001),
                          pressure = c("GHG", "nutrients", "nutrients", "water", "disturbance"))

pressures_converted <- left_join(pressures, conversions, by = "stressor") %>%
  mutate(pressure_per_tonne = median*conversion) %>%
  select(stressor_group, pressure, pressure_per_tonne) %>%
  group_by(stressor_group, pressure) %>%
  summarize(pressure_per_tonne = sum(pressure_per_tonne)) %>%
  ungroup()

pressures_lw <- left_join(tonnes_lw, pressures_converted, by="stressor_group") %>%
  mutate(stressor_value = tonnes_lw*pressure_per_tonne) 

pressures_lw %>%
  group_by(stressor_group, pressure) %>%
  summarize(stressor_value = sum(stressor_value)) %>%
  data.frame()

write_csv(pressures_lw, here("scripts/fw_aquaculture_pressures/output/fw_aquaculture_pressures.csv"))
```

## Crop feed composition

- Estimate proportion of oilcrops vs. grains
Mostly from food_systems/feed/diet_composition_wrangling.xlsx
For catfish: mostly: https://agric4profits.com/the-ideal-feeds-for-catfish/
https://thefishsite.com/articles/catfish-nutrition-feeds-and-feed-formulation
https://www.fao.org/3/Q3567E/q3567e05.htm
rice bran: https://onlinelibrary.wiley.com/doi/10.1111/J.1365-2109.2011.03048.X
rice bran: https://fishfeedmachinery.com/Solution/nutritional-fish-feed-formulation.html

For tilapia: https://www.fao.org/fishery/affris/species-profiles/nile-tilapia/feed-formulation/en/

- Ignores animal products (beyond fish oil fish meal)

- Use fofm data from Halley for each group and then distribute based on tonnes of production in each country

- rescale hypoph carps so feed sums to 1


```{r}

taxa_list <- unique(tonnes_lw$stressor_group)
crop_cats <- read_csv(here("scripts/fw_aquaculture_pressures/inputs/crop_breakout.csv"))
list.files("/home/shares/food-systems/social_justice_projects/_raw_data/fresh_water_aqua_pressures_Jessica/")

feed <- read_csv("/home/shares/food-systems/social_justice_projects/_raw_data/fresh_water_aqua_pressures_Jessica/aquaculture_group_fcr_feed_averages.csv") %>%
  filter(taxa %in% taxa_list) %>%
  rowwise() %>%
  mutate(total = feed_soy + feed_crops + feed_fmfo + feed_animal) %>%
  mutate(feed_soy = feed_soy/total,
         feed_crops = feed_crops/total,
         feed_fmfo = feed_fmfo/total,
         feed_animal = feed_animal/total) %>%
  left_join(crop_cats, by="taxa") %>%
  mutate(feed_grain = feed_crops * grains) %>%
  mutate(feed_oil = feed_crops * oil_crops) %>%
  mutate(total = feed_soy + feed_grain + feed_oil + feed_fmfo + feed_animal) %>%
  dplyr::select(stressor_group = taxa, fcr, feed_soy,feed_grain, feed_oil, feed_fmfo, total)

```

## Crop feed quantity
Calculate tonnes of each feed category for each country and taxa group based on fcr, diet composition, and tonnes of production.

```{r}

feed_tonnes_pre <- left_join(tonnes_lw, feed, by="stressor_group") %>%
  rowwise() %>%
  mutate(tonnes_feed_soy = tonnes_lw * fcr * feed_soy) %>%
  mutate(tonnes_feed_grain = tonnes_lw * fcr * feed_grain) %>%
  mutate(tonnes_feed_oil = tonnes_lw * fcr * feed_oil) %>%
  select(iso3c, country_name, stressor_group, tonnes_feed_soy, tonnes_feed_grain, tonnes_feed_oil) %>%
  pivot_longer(c("tonnes_feed_soy", "tonnes_feed_grain", "tonnes_feed_oil"), names_to = "feed_type")

```

Correct for averaged loss during processing. This is based on averaged data for each feed category in food_systems/feed/data/feed_extraction_rates.csv. This is a very rough estimate.

```{r}

crop_loss <- data.frame(feed_type = c("tonnes_feed_soy", "tonnes_feed_grain", "tonnes_feed_oil"), 
                        loss_percent = c(3, 2, 9))

crop_feed_tonnes <- left_join(feed_tonnes_pre, crop_loss, by="feed_type") %>%
    mutate(tonnes_product = value * 100/(100 - loss_percent)) %>%
    mutate(feed_category = ifelse(feed_type == "tonnes_feed_soy", "soy", NA)) %>%
    mutate(feed_category = ifelse(feed_type == "tonnes_feed_grain", "grain", feed_category)) %>%
    mutate(feed_category = ifelse(feed_type == "tonnes_feed_oil", "oil", feed_category)) %>%
  select(iso3c, country_name, taxa=stressor_group, feed_category, tonnes=tonnes_product)

write_csv(crop_feed_tonnes, here("scripts/fw_aquaculture_pressures/output/fw_aqua_crop_feed.csv"))
```

## fishoil fish meal consumption
I am going to use the values from Halley et al. and distribute by production.

We will be using the values from Halley as the baseline, and then distribute consumption of fofm based on production.
```{r}

fofm_reference <- read_csv(here("scripts/fw_aquaculture_pressures/inputs/msleckman.45.1_fofm_consumption.csv")) %>%
  filter(!is.na(fw_aquaculture))

## need to make a few adjusments to divide categories and such
# 1. We use catfish pressures for ISSCAAP "Miscellaneous freshwater fishes". These are mostly catfish, and difficult to identify.
#    But there is reported aquaculture for: "Miscellaneous coastal fishes" and "Marine fishes not identified", but these end up  being <<1% of total production so we can ignore these contributions.

fw_fishes_not_inc <- filter(production_update, is.na(stressor_group)) %>%
  filter(ISSCAAP_Group_En %in% c("Miscellaneous coastal fishes", "Marine fishes not identified"))
fw_fishes_inc <- filter(production_update, ISSCAAP_Group_En %in% c("Miscellaneous coastal fishes", "Marine fishes not identified", "Miscellaneous freshwater fishes"))
sum(fw_fishes_not_inc$tonnes_lw)/sum(fw_fishes_inc$tonnes_lw) # <<1% can assume all fofm goes to catfish

# 2. fofm reference includes diadromous fishes, which the pressures break into two categories:
# milkfish and misc_diads (Barramundi and a very small amount of striped bass)

tmp <- filter(production_update, ISSCAAP_Group_En %in% "Miscellaneous diadromous fishes") %>% data.frame()
milkfish <- filter(production_update, stressor_group == "milkfish")
other_diad <- filter(production_update, stressor_group == "misc_diad")
sum(milkfish$tonnes_lw)/(sum(milkfish$tonnes_lw)+sum(other_diad$tonnes_lw)) # 95% milkfish!
# but need to take into account proportion fofm and fcr in diet from Jessica's data:
(sum(milkfish$tonnes_lw)*0.0571*1.68)/((sum(milkfish$tonnes_lw)*0.0571*1.68) + (sum(other_diad$tonnes_lw) * 0.304*1.34)) # 81% fofm by milkfish!

# 3. Carps were divided into hypoph_carp and other carp
hypoph_carp <- filter(production_update, stressor_group == "hypoph_carp")
oth_carp <- filter(production_update, stressor_group == "oth_carp")
sum(hypoph_carp$tonnes_lw)/(sum(hypoph_carp$tonnes_lw)+sum(oth_carp$tonnes_lw)) # 28% hypoph_carp
(sum(hypoph_carp$tonnes_lw) * 0.0102 * 0.905)/(sum(hypoph_carp$tonnes_lw)*0.0102 * 0.905 + (sum(oth_carp$tonnes_lw)*0.0531*1.78)) # 4% fofm to hypoph carp

#4. need a salmon correction in main data 
698684/sum(c(13930, 2101278, 698684)) # multiply marine salmon by 0.75, given that we only included marine, brackish salmon 

# this also generaly matches production
raw_production <- read_csv("/home/shares/food-systems/social_justice_projects/_raw_data/fao_aquaculture/Aquaculture_Quantity.csv") %>%
  rename(species_code = "SPECIES.ALPHA_3_CODE") %>%
  left_join(species, by="species_code") %>%
  filter(ISSCAAP_Group_En == "Salmons, trouts, smelts") %>%
  group_by(ENVIRONMENT.ALPHA_2_CODE) %>%
  summarize(tonnes = sum(VALUE))


divided_fofm <- data.frame(fw_aquaculture=c("milkfish", "misc_diad", "hypoph_carp", "oth_carp"),
                           prop =c(0.81, 0.19, 0.04, 0.96))

fofm <- left_join(fofm_reference, divided_fofm, by="fw_aquaculture") %>%
  mutate(prop = ifelse(is.na(prop), 1, prop)) %>%
  mutate(taxa_tonnes_fofm_consumption = Current_Mean*prop) %>%
  select(stressor_group = fw_aquaculture, taxa_tonnes_fofm_consumption)

```

Distribute fofm consumption based on production by country

```{r}
fofm_tonnes <- tonnes_lw %>%
  group_by(stressor_group) %>%
  mutate(global_tonnes=sum(tonnes_lw)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(prop_prod = tonnes_lw/global_tonnes) %>%
  left_join(fofm, by="stressor_group") %>%
  mutate(tonnes_fofm = taxa_tonnes_fofm_consumption*prop_prod) %>%
  select(iso3c, country_name, taxa=stressor_group, tonnes_fofm)  

fofm_tonnes %>% 
  group_by(taxa) %>%
  summarize(total=sum(tonnes_fofm))
  
write_csv(fofm_tonnes, here("scripts/fw_aquaculture_pressures/output/fw_fofm_consumption.csv")) 
```