---
title: "connect_2017_fw_aquaculture_to_pressures"
author: "Haley Epperly"
date: "2022-11-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```
## Set Up

File paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
aquatic_trade_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01c_allocate_unknown_trade_fish_consumption/"
pressures_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/02b_fw_aquaculture_pressures/"
code_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/Codes/"
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/04a_connect_fw_aquaculture_pressures/"
```

Method:
The trade data contains the source country, consumer country, habitat, fishing method, nceas group, and the live weight traded between each country. The pressures data has the environmental pressure produced per tonne live weight of a given species. The pressures data is a global average for freshwater aquaculture so we only need to match the species traded with each pressure and multiply the two to get the total environmental pressure for the consuming country.

## Read in the two data sources
Trade data - prepared in step 01c
```{r}
# read in all of the fish trade data, these are means of the years 2015-2019
trade_matrix_raw <- read_csv(paste0(aquatic_trade_data_path, "fish_trade_matrix_w_unknowns_allocated.csv")) # read in the updated data

# only want inland aquaculture to accurately allocate freshwater (inland) aquaculture pressures
fw_trade_clean <- trade_matrix_raw %>% 
  filter(habitat == "inland" & method == "aquaculture")
```

Read in freshwater aquaculture pressures data per tonne from Mel - prepared in step 02a
```{r}
pressure <- read_csv(paste0(pressures_path, "fw_aquaculture_pressures_per_tonne.csv")) 
```

Set up country codes to join country names
```{r}
# read in country codes from FAO detailed trade matrix definitions selecting the partner country group tab (reported country group tab says data unavailable)
country_codes <- read_csv(paste0(code_path, "FAOSTAT_country_codes_8-10-2022.csv"))

# format to match trade matrix variables names
countries <- country_codes %>% 
  select(`Partner Country Code`, `Partner Countries`, `ISO3 Code`) %>% 
  rename('Country_Code' = 'Partner Country Code',
         'Country' = 'Partner Countries',
         'ISO3' = 'ISO3 Code') %>% 
  unique()

# "China, mainland" on the FAO website has the ISO3 code F41.
# China, the aggregated country, has the ISO3 code CHN.
# In the pressures data, they assign China mainland the ISO3 code CHN, so we need to be aware
# of this in the gravity model if we match using ISO3 code. 
# we will keep in line with the pressures data here, and give China mainland the ISO3 code CHN.
countries <- countries %>% 
  mutate(ISO3 = ifelse(Country == "China, mainland", "CHN", ISO3))

```

## Match up the species groupings
Compare species groups - these are extremely different. We will need to make sure that we can join the pressures per tonne of each species to their correct species in the trade data
```{r}
table(pressure$stressor_group)
table(fw_trade_clean$nceas_group)
```

Investigating salmon to make sure Salmons, trouts, smelts are all inland aquaculture and don't include marine species.
```{r}
salmonid <- trade_matrix_raw %>% 
  filter(nceas_group=="salmonids")
salmon <- trade_matrix_raw %>% 
  filter(nceas_group=="salmons, trouts, smelts")
```
Salmonids are all marine aquaculture, so Salmons, trouts, smelts `nceas_group` is probably just inland.

Fish matching key - updated data (left is trade matrix categories, right is pressures categories):
oth_carp = oth_carp 
hypoph_carp = hypoph_carp
miscellaneous diadromous fishes = misc_diad 
salmons, trouts, smelts = trout
miscellaneous freshwater fishes = catfish
tilapias and other cichlids = tilapia

Re-label pressures data to categories used in trade data (pressures data is more disaggregated)
```{r}
pressure_relabel <- pressure %>% 
  mutate(nceas_group = case_when(stressor_group == "catfish" ~ "miscellaneous freshwater fishes",
                                 stressor_group == "hypoph_carp" ~ "hypoph_carp",
                                 stressor_group == "misc_diad" ~ "miscellaneous diadromous fishes",
                                 stressor_group == "oth_carp" ~ "oth_carp",
                                 stressor_group == "tilapia" ~ "tilapias and other cichlids",
                                 stressor_group == "trout" ~ "salmons, trouts, smelts",
                                 TRUE ~ as.character(NA)))
summary(pressure_relabel)

pressure_relabel <- pressure_relabel %>% 
  mutate(pressure = case_when(pressure == "GHG" ~ "ghg",
                              pressure == "nutrients" ~ "nutrient",
                              TRUE ~ pressure)) %>% 
  filter(!is.na(nceas_group)) %>% 
  select(nceas_group, pressure, pressure_per_tonne, stressor_group) %>% 
  pivot_wider(names_from = pressure,
              values_from = pressure_per_tonne)

```

Subset trade to relevant categories for which we have pressures data (removes ~ 25% of rows). There is not freshwater aquaculture pressures for the following groups:
- freshwater crustaceans
- sturgeons, paddlefishes
- river eels
- miscellaneous coastal fishes
- marine fishes not identified
- freshwater molluscs
- shads
- miscellaneous marine crustaceans
- shrimps, prawns
- abalones, winkles, conchs

```{r}
fw_trade_subset <- fw_trade_clean %>% 
  filter(nceas_group == "oth_carp" | 
           nceas_group == "hypoph_carp" |
           nceas_group == "miscellaneous diadromous fishes" |
           nceas_group == "salmons, trouts, smelts" |
           nceas_group == "miscellaneous freshwater fishes"|
           nceas_group == "tilapias and other cichlids")

# check to see how much of trade data is accounted for when subset by available pressure data
nrow(fw_trade_subset)/nrow(fw_trade_clean)
# 75% of trade by rows

#check by weight how much trade is unaccounted for
sum(fw_trade_subset$consumption_live_weight_t)/sum(fw_trade_clean$consumption_live_weight_t)
# more than 92% of the trade by weight
```

## Calculate the environmental pressures from freshwater aquaculture
Now join trade data with pressure efficiency data by the `nceas_group`. The `fw_trade_pressures` should have the same amount of rows as `fw_trade_subset`.
```{r}
# join trade data with the pressures data
fw_trade_pressures <- left_join(x = fw_trade_subset, y = pressure_relabel, by = "nceas_group")

fw_trade_pressures <- fw_trade_pressures %>% 
  # calculate the total pressure traded by multiplying consumption_live_weight_t by the pressure efficiency
  mutate(disturbance_total = disturbance*consumption_live_weight_t,
         ghg_total = ghg*consumption_live_weight_t,
         nutrient_total = nutrient*consumption_live_weight_t,
         water_total = water*consumption_live_weight_t) 
```

## Set up for joining with trade and pressures data and save
```{r}
fw_trade_pressures_clean <- fw_trade_pressures %>% 
  # filter out NEI iso3c code
  filter(consumer_iso3c != "NEI",
         source_country_iso3c != "NEI") %>% 
  # join with countries data to producer iso3c
  left_join(countries %>% 
              select(-Country_Code),
            by = c("source_country_iso3c" = "ISO3")) %>%
  # rename to producer columns
  rename(producer_country = Country, producer_iso3c = source_country_iso3c) %>% 
  # relocate to after producer country code
  relocate(c(producer_country), .after = producer_iso3c) %>% 
  # join with countries data to consumer iso3c
  left_join(countries %>% 
              select(-Country_Code),
            by = c("consumer_iso3c" = "ISO3")) %>%
  # rename to producer columns
  rename(consumer_country = Country, consumer_iso3c = consumer_iso3c) %>% 
  # relocate to after producer country code
  relocate(c(consumer_country), .after = consumer_iso3c) %>% 
  # make sure food group is the same for all categories
  mutate(food_group = "fisheries and aquaculture",
         # add in item code column for fisheries and aquaculture, will be NA
         item_code = NA) %>% 
  # add item column for freshwater aquaculture
  mutate(item = case_when(nceas_group == "miscellaneous freshwater fishes" ~ "fw_aquaculture_miscellaneous_freshwater_fishes",
                          nceas_group == "hypoph_carp" ~ "fw_aquaculture_hypoph_carp",
                          nceas_group == "miscellaneous diadromous fishes" ~ "fw_aquaculture_miscellaneous_diadromous_fishes",
                          nceas_group == "oth_carp" ~ "fw_aquaculture_oth_carp",
                          nceas_group == "tilapias and other cichlids" ~ "fw_aquaculture_tilapias_and_other_cichlids",
                          nceas_group == "salmons, trouts, smelts" ~ "fw_aquaculture_salmons_trouts_smelts",
                          TRUE ~ NA)) %>% 
  # rename consumption_live_weight_t
  rename(traded_tonnes = consumption_live_weight_t) %>% 
  #excess columns
  select(-c(ghg, disturbance, nutrient, water, nceas_group, habitat, method, stressor_group))

```

## Save
```{r}
write.csv(x = fw_trade_pressures_clean,
          file = paste0(save_path, "fw_aquaculture_trade_pressures.csv"))
```

