---
title: "Calculating proportion crop used as feed using 2017 FAO food balance data"
author: "Haley Epperly"
date: "2022-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

Objective: Calculate the proportion of each country's domestic supply quantity that they used as feed for each relevant crop using 2017 FAO Food Balance data.

Check: USDA says the USA uses ~40% of their corn for feed, which is in line with calculations using 2017 FAO FB data and dividing feed by domestic supply quantity for maize (41.6%).
Source: https://www.ers.usda.gov/topics/crops/corn-and-other-feedgrains/feedgrains-sector-at-a-glance/#:~:text=Corn%20is%20a%20major%20component,of%20total%20domestic%20corn%20use.

Steps:
1. Download FAO FB data subset to 2017
2. For each crop and country, divide the tonnes in feed by tonnes domestic supply quantity
3. Combine dfs for all grains
4. Download FAO FB data with regions
5. Merge grains df with region by country name (countrycode package)
6. Gapfill missing data with region data
7. Save grains output
8. Repeat for pulses

Grains (and name in FAO FB data [2010-]):
- barl - barley (FAO = Barley and products)
- maiz -  maize (FAO = Maize and products)
- sorg- sorghum (FAO = Sorghum and products)
- xmil - millet (FAO = Millet and products)
- rice - rice (FAO = Rice and products)
- whea - wheat (FAO = Wheat and products)
- ocer (other cereals) includes FAO categories 1. Rye and Products, 2. Oats, and 3. Cereals, Other (Default composition: 68 Popcorn, 89 Buckwheat, 90 Flour, buckwheat, 91 Bran, buckwheat, 92 Quinoa, 94 Fonio, 95 Flour, fonio, 96 Bran, fonio, 97 Triticale, 98 Flour, triticale, 99 Bran, triticale, 101 Canary seed, 103 Grain, mixed, 104 Flour, mixed grain, 105 Bran, mixed grains, 108 Cereals, nes, 111 Flour, cereals, 112 Bran, cereals nes, 113 Cereal preparations, nes)

Pulses (xpul - one MapSPAM category):
- Beans
- Peas
- Pulses, Other and products (Default composition: 181 Broad beans, horse beans, dry, 191 Chick peas, 195 Cow peas, dry, 197 Pigeon peas, 201 Lentils, 203 Bambara beans, 205 Vetches, 210 Lupins, 211 Pulses, nes, 212 Flour, pulses, 213 Bran, pulses)

Notes about other crops:
- sugb (sugar beets) - 0 tonnes (out of 31,096,000 tonnes) in US go to feed. Across the world, 14,405,000/304,420,000 tonnes (5%) go to feed. Sugar is also quite complicated because it combines sugar beet and cane, so probably want to not factor in feed.
- cass (cassava and products) 86,615,000/281,449,000 tonnes (31%) in world go to feed. Maybe include this?
- bana (bananas) 1,011,000/112,069,000 tonnes (0.9%) in world go to feed, so don't include this because it's such a small amount.


MapSPAM to FAO - to compare categories you can look at this df
https://github.com/OHI-Science/food_systems/blob/master/feed/data/MapSPAM_to_FAO_v2.csv

## Set Up
Load libraries
```{r}
library(tidyverse)
library(countrycode)
```

Data Paths
```{r}
raw_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
FAO_path <- paste0(raw_path, "FAO/")
kastner_path <- paste0(raw_path, "kastner_2022/")
fish_trade_path <- paste0(raw_path, "trade/aquatic_food_trade/")
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/03a_calc_proportion_feed_per_crop_country/"
```


Read in FAO FB data (2010-)
Regions are not included in this df, only countries. To gapfill, will also need to calculate these values for regions.
```{r}
fb <- read_csv(paste0(kastner_path, "CommodityBalances_Crops_2017_E_All_Data_(Normalized)_7_20_22.csv"))

# only 2017 data? - Yes
range(fb$Year)

# clean up data
fb <- fb %>% 
  select(`Area Code (FAO)`, Area, Element, `Item Code`, Item, Unit, Value) %>% 
  filter(Element == "Domestic supply quantity" | Element == "Feed") %>% 
  # converting to tonnes (instead of 1000 tonnes) in case we want to see raw tonnes used for feed later
  mutate(value_tonnes = Value*1000) %>% 
  select(-c(Unit, Value))
```


Create function to calculate proportion feed for each category
```{r}
calc_prop_feed <- function(FAO_name, feed_prop_col_name, dsq_col_name, feed_col_name) {
  FAO_name <- sym(FAO_name)
  fb %>% 
    filter(Item == FAO_name) %>% 
    select(-`Item Code`) %>% 
    pivot_wider(names_from = Element, values_from = value_tonnes) %>% 
    mutate({{feed_prop_col_name}} := Feed/`Domestic supply quantity`) %>% 
    rename({{dsq_col_name}} := `Domestic supply quantity`) %>% 
    rename({{feed_col_name}} := Feed) %>% 
    select(-Item)
}
```


### For Gapfilling

We need to add in missing countries that are missing from FB and CB data. We will use a combination of FAO country names and additional countries from fisheries trade data. We won't need all of these countries, but it's good to include them all just in case.
```{r}
# rather than reading in the whole FAO country list, we'll just consider countries included in the Kastner trade matrices output
# read in FAO country list
tm_raw <- read_csv(paste0(kastner_path, "trade_matrix_by_country_of_origin_2017_import.csv"))

tm <- tm_raw %>% 
  # just select the country codes
  select(Consumer_Country_Code) %>% 
  unique() %>% 
  # match the country code to the country names
  mutate(country = countrycode(Consumer_Country_Code, "fao", "country.name")) %>% 
  # change some of the names for our needs
  mutate(country = case_when(Consumer_Country_Code == 41 ~ "China, mainland",
                             Consumer_Country_Code == 299 ~ "Palestine",
                             TRUE ~ country))

# read in fisheries trade data 
fish_tm_raw <- read_csv(paste0(fish_trade_path, "20230626_nceas_consumption.csv"))

# add country names and subset
fish_tm <- fish_tm_raw %>% 
  # just select the country codes
  select(consumer_iso3c) %>% 
  unique() %>% 
  # match the country code to the country names
  mutate(country = countrycode(consumer_iso3c, 'iso3c', 'country.name')) %>% 
  # change some of the names for our needs
  filter(!is.na(country)) # remove one NA (iso3c = NEI, Netherlands East Indies, not in FAO data either)


# combine dfs to get complete list of countries
countries <- full_join(tm, fish_tm, by = "country")

# add iso3c
countries <- countries %>% 
  mutate(iso3 = countrycode(country, "country.name", "iso3c")) %>% 
  select(country, iso3)
```

## Grains
Run through function with all relevant grains
- Barley and products
- Maize and products
- Sorghum and products
- Millet and products
- Rice and products
- Wheat and products
- Rye and Products
- Oats
- Cereals, Other 
```{r}
barley <- calc_prop_feed("Barley and products", "barl_prop_feed", "barl_dsq", "barl_feed")
maize <- calc_prop_feed("Maize and products", "maiz_prop_feed", "maiz_dsq", "maiz_feed")
sorghum <- calc_prop_feed("Sorghum and products", "sorg_prop_feed", "sorg_dsq", "sorg_feed")
millet <- calc_prop_feed("Millet and products", "mill_prop_feed", "mill_dsq", "mill_feed")
rice <- calc_prop_feed("Rice and products", "rice_prop_feed", "rice_dsq", "rice_feed")
wheat <- calc_prop_feed("Wheat and products", "whea_prop_feed", "whea_dsq", "whea_feed")
rye <- calc_prop_feed("Rye and products", "rye_prop_feed", "rye_dsq", "rye_feed")
oats <- calc_prop_feed("Oats", "oats_prop_feed", "oats_dsq", "oats_feed")
other_cereal <- calc_prop_feed("Cereals, Other", "ocer_prop_feed", "ocer_dsq", "ocer_feed")

```

Some feed proportions will need to be gapfilled depending on what the joint data shows as they either have either NA, NaN, or zero values.

### Calculate proportion feed for all grains
Bind grains into one df
```{r}
grains <- barley %>% 
  full_join(maize, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(sorghum, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(millet, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(rice, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(wheat, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(rye, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(oats, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(other_cereal, by = c("Area Code (FAO)", "Area"))

# calculate an average percent feed for grains by summing all feed and dividing by sum of all dsq
grains <- grains %>% 
  rowwise() %>% 
  mutate(total_dsq = sum(barl_dsq, maiz_dsq, sorg_dsq, mill_dsq, rice_dsq, whea_dsq, rye_dsq, oats_dsq, ocer_dsq, na.rm = TRUE)) %>% 
  mutate(total_feed = sum(barl_feed, maiz_feed, sorg_feed, mill_feed, rice_feed, whea_feed, rye_feed, oats_feed, ocer_feed, na.rm = TRUE)) %>% 
  mutate(total_prop_feed = total_feed/total_dsq) %>% 
  select(`Area Code (FAO)`, Area, total_dsq, total_feed, total_prop_feed)

# save grains csv before gapfilling
#write_csv(grains, paste0(save_path, "perc_dsq_grains_used_as_feed.csv"))
```


### Gapfill Grains
```{r}
region_fb <- read_csv(paste0(FAO_path, "2017_FAO_data/FAO_Food_Balances_2017_grains_pulses_feed_dl_11_2_22.csv"))

# all 2017? - yes
range(region_fb$Year)

grain_region_fb <- region_fb %>% 
  # FB data comes in units of 1000 tonnes so convert to tonnes
  mutate(value_tonnes = Value*1000) %>% 
  # remove extra columns
  select(-c(`Element Code`, `Item Code (CPC)`, `Year Code`, Year, Unit, Value)) %>% 
  pivot_wider(names_from = Element, values_from = value_tonnes) %>% 
  # calculate proportion of grain used for feed
  mutate(prop_feed = Feed/`Domestic supply quantity`) %>% 
  # remove beans, pulses, and peas
  filter(!Item == "Beans", !Item == "Pulses, Other and products", !Item == "Peas") %>% 
  # group by area and sum feed and dsq of grains
  group_by(Area) %>% 
  summarise(sum_dsq = sum(`Domestic supply quantity`, na.rm=TRUE), 
            sum_feed = sum(Feed, na.rm=TRUE)) %>% 
  # calculate aggregate grain proportion
  mutate(total_prop_feed = sum_feed/sum_dsq)

```

Based on `grain_region_fb` the only region that has no feed data is Micronesia. As Polynesia and Melanesia are the closest regions to Micronesia geographically and culturally we will use an average of their feed proportions for Micronesia.
```{r}
grain_region_fb$total_prop_feed[grain_region_fb$Area == "Micronesia"] <- mean(grain_region_fb$total_prop_feed[grain_region_fb$Area == "Polynesia"], grain_region_fb$total_prop_feed[grain_region_fb$Area == "Melanesia"])
```


Merge grains df with list of countries from trade matrices to ensure all countries are gapfilled even if there wasn't any FAO FB or CB data for them. Add in the subregion and intermediate regions for the countries
```{r}
# add iso3 to grains
grains_countries<- grains %>% 
  mutate(iso3 = countrycode(Area, "country.name", "iso3c")) %>% 
  # Turkey comes up as NA for iso3c because of lettering, fill this in
  mutate(iso3 = case_when(is.na(iso3) ~ "TUR", TRUE ~ iso3)) %>% 
  # join with country list
  full_join(countries, by = "iso3") %>% 
  # fill in country name where NA
  mutate(Area = case_when(is.na(Area) ~ country, TRUE ~ Area)) %>% 
  select(-country) %>% 
  # this gets rid of duplicates of China and China, Mainland
  unique() %>% 
  # remove Palestinian Territories because we have Palestine
  filter(!Area == "Palestinian Territories")

# Add in the regional information to grains_countries
grains_countries <- grains_countries %>% 
  # add subregion and intermediate region
  mutate(subregion = countrycode(iso3, origin = 'iso3c', destination = 'un.regionsub.name'),
         intermed_region = countrycode(iso3, origin = 'iso3c', destination = 'un.regionintermediate.name')) %>% 
  # make region column where the intermediate region is used unless it is missing, then use subregion
  mutate(region = case_when(is.na(intermed_region) ~ subregion,
                            TRUE ~ intermed_region)) %>% 
  # reorder columns 
  select(Area, iso3, region, total_dsq, total_feed, total_prop_feed)

# fill in Taiwan 
grains_countries$region[grains_countries$iso3 == "TWN"] <- "Eastern Asia"
grains_countries$Area[grains_countries$iso3 == "TWN"] <- "Taiwan"

```

Now gapfill using the regional data!
```{r}
# select only region and regions_prop_feed to join with country data
grain_region_fb <- grain_region_fb %>% 
  select(region = Area, region_prop_feed = total_prop_feed)


grains_gap_filled <- left_join(grains_countries, grain_region_fb) %>% 
  # gapfill when total_prop_feed is 0 or NA, otherwise use regional average
  mutate(grain_prop_feed_gf = case_when(total_prop_feed == 0 ~ region_prop_feed,
                                        is.na(total_prop_feed) ~ region_prop_feed,
                                        TRUE ~ total_prop_feed)) %>% 
  # select only the columns we need
  select(Area, iso3, region, grain_prop_feed_gf)

write_csv(grains_gap_filled, paste0(save_path, "perc_dsq_grains_used_as_feed_gapfilled.csv"))

```


## Pulses

For pulses, need to combine all tonnes and then calculate proportions to make sure it's weighted correctly since we only have one xpul category for MapSPAM.

```{r}
# format beans
beans <- fb %>% 
  filter(Item == "Beans") %>% 
  select(-c(`Item Code`, Item)) %>% 
  pivot_wider(names_from = Element, values_from = value_tonnes) %>% 
  rename(beans_dsq = `Domestic supply quantity`,
         beans_feed = Feed)

# format peas
peas <- fb %>% 
  filter(Item == "Peas") %>% 
  select(-c(`Item Code`, Item)) %>% 
  pivot_wider(names_from = Element, values_from = value_tonnes) %>% 
  rename(peas_dsq = `Domestic supply quantity`,
         peas_feed = Feed)

# format Pulses, Other and Products
other_pulse <- fb %>% 
  filter(Item == "Pulses, Other and products") %>% 
  select(-c(`Item Code`, Item)) %>% 
  pivot_wider(names_from = Element, values_from = value_tonnes) %>% 
  rename(other_pulse_dsq = `Domestic supply quantity`,
         other_pulse_feed = Feed)

# bind together
pulses <- beans %>% 
  full_join(peas, by = c("Area Code (FAO)", "Area")) %>% 
  full_join(other_pulse, by = c("Area Code (FAO)", "Area"))
```

### Calculate proportion of pulses for feed
Combine dsq and feed for each crop, then calculate proportion feed from total DSQ. If pulse_dsq = 0, then prop_feed = NaN, otherwise prop_feed will = 0 if no values (0s or NAs) in pulse_feed.
```{r}
pulses <- pulses %>%
  # have to use rowwise - otherwise it sums the whole column, similar to groupby(Area) 
  rowwise() %>% 
  mutate(pulse_dsq = sum(beans_dsq, peas_dsq, other_pulse_dsq, na.rm=TRUE)) %>% 
  mutate(pulse_feed = sum(beans_feed, peas_feed, other_pulse_feed, na.rm=TRUE)) %>% 
  mutate(total_prop_feed = pulse_feed/pulse_dsq)

# save pulses df before gapfilling
#write_csv(pulses, paste0(save_path, "perc_dsq_pulses_used_as_feed.csv"))
```

### Gapfill Pulses
```{r}
region_fb <- read_csv(paste0(FAO_path, "2017_FAO_data/FAO_Food_Balances_2017_grains_pulses_feed_dl_11_2_22.csv"))

# all 2017? - yes
range(region_fb$Year)

pulse_region_fb <- region_fb %>% 
  # FB data comes in units of 1000 tonnes so convert to tonnes
  mutate(value_tonnes = Value*1000) %>% 
  # remove extra columns
  select(-c(`Element Code`, `Item Code (CPC)`, `Year Code`, Year, Unit, Value)) %>% 
  pivot_wider(names_from = Element, values_from = value_tonnes) %>% 
  # calculate proportion of Item used for feed
  mutate(prop_feed = Feed/`Domestic supply quantity`) %>% 
  # filter for beans, pulses, and peas
  filter(Item %in% c("Beans", "Pulses, Other and products", "Peas")) %>% 
  # group by area and sum feed and dsq of pulses
  group_by(Area) %>% 
  summarise(sum_dsq = sum(`Domestic supply quantity`, na.rm=TRUE), 
            sum_feed = sum(Feed, na.rm=TRUE)) %>% 
  # calculate aggregate pulses proportion
  mutate(total_prop_feed = sum_feed/sum_dsq)

```

Based on `pulse_region_fb` the only regions that have no feed data are Micronesia, Polynesia, and Melanesia. As New Zealand and Australia make up the rest of the Oceania region and their proportions are the higher end of the spectrum it would be more conservative to use the world average rather than the regional average. Also the economies of New Zealand and Australia are quite different than Micronesia, Polynesia, and Melanesia.

```{r}
pulse_region_fb$total_prop_feed[pulse_region_fb$Area == "Micronesia"] <- pulse_region_fb$total_prop_feed[pulse_region_fb$Area == "World"]
pulse_region_fb$total_prop_feed[pulse_region_fb$Area == "Polynesia"] <- pulse_region_fb$total_prop_feed[pulse_region_fb$Area == "World"]
pulse_region_fb$total_prop_feed[pulse_region_fb$Area == "Melanesia"] <- pulse_region_fb$total_prop_feed[pulse_region_fb$Area == "World"]

```


Merge pulses df with list of countries from trade matrices to ensure all countries are gapfilled even if there wasn't any FAO FB or CB data for them. Add in the subregion and intermediate regions for the countries
```{r}
# add iso3 to pulses
pulses_countries<- pulses %>% 
  mutate(iso3 = countrycode(Area, "country.name", "iso3c")) %>% 
  # Turkey comes up as NA for iso3c because of lettering, fill this in
  mutate(iso3 = case_when(is.na(iso3) ~ "TUR", TRUE ~ iso3)) %>% 
  # join with country list
  full_join(countries, by = "iso3") %>% 
  # fill in country name where NA
  mutate(Area = case_when(is.na(Area) ~ country, TRUE ~ Area)) %>% 
  select(-country) %>% 
  # this gets rid of duplicates of China and China, Mainland
  unique() %>% 
  # remove Palestinian Territories because we have Palestine
  filter(!Area == "Palestinian Territories")

# Add in the regional information to pulses_countries
pulses_countries <- pulses_countries %>% 
  # add subregion and intermediate region
  mutate(subregion = countrycode(iso3, origin = 'iso3c', destination = 'un.regionsub.name'),
         intermed_region = countrycode(iso3, origin = 'iso3c', destination = 'un.regionintermediate.name')) %>% 
  # make region column where the intermediate region is used unless it is missing, then use subregion
  mutate(region = case_when(is.na(intermed_region) ~ subregion,
                            TRUE ~ intermed_region)) %>% 
  # reorder columns 
  select(Area, iso3, region, pulse_dsq, pulse_feed, total_prop_feed)

# fill in Taiwan 
pulses_countries$region[pulses_countries$iso3 == "TWN"] <- "Eastern Asia"
pulses_countries$Area[pulses_countries$iso3 == "TWN"] <- "Taiwan"

```

Now gapfill using the regional data!
```{r}
# select only region and regions_prop_feed to join with country data
pulse_region_fb <- pulse_region_fb %>% 
  select(region = Area, region_prop_feed = total_prop_feed)


pulses_gap_filled <- left_join(pulses_countries, pulse_region_fb) %>% 
  # gapfill when total_prop_feed is -1, 0,  or NA using regional average. If > 1 make proportion 1 
  mutate(pulses_prop_feed_gf = case_when(total_prop_feed <= 0 ~ region_prop_feed,
                                         is.na(total_prop_feed) ~ region_prop_feed,
                                         total_prop_feed > 1 ~ 1,
                                         TRUE ~ total_prop_feed)) %>% 
  # select only the columns we need
  select(Area, iso3, region, pulses_prop_feed_gf)

write_csv(pulses_gap_filled, paste0(save_path, "perc_dsq_pulses_used_as_feed_gapfilled.csv"))

```



