---
title: "02ai_fw_aquaculture_pressures"
output: html_document
date: "2022-11-02"
editor_options: 
  chunk_output_type: console
---

GOAL: Calculate crop and FOFM consumption by freshwater aquaculture by country and taxa group.
```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(countrycode)
```

File Paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
fao_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/FAO/fao_aquaculture/"
fw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/fw_aquaculture/"
aquatic_trade_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01c_allocate_unknown_trade_fish_consumption/"
pressures_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/pressures/"
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/03b_fofm_fw_aquaculture_feed/"
```


## FAO production data - Freshwater Aquaculture
To calculate the total amount of crops needed for freshwater aquaculture feed, we will use FAO production data for freshwater aquaculture.

Format these data to include only the years of interest
IN = freshwater
BW = brackishwater
MA = marine
Al = all environments
```{r}
# read in production data
raw_production <- read_csv(paste0(fao_data_path, "Aquaculture_Quantity.csv")) 

production_clean <- raw_production %>%
  # filter for freshwater aquaculture
  filter(ENVIRONMENT.ALPHA_2_CODE=="IN") %>%
  # filter for 2015-2019 production years to take average, could just do 2017 to match pressures data
  filter(PERIOD %in% c(2015, 2016, 2017, 2018, 2019)) %>% 
  # select columns of interest
  select(UN_Code = COUNTRY.UN_CODE, species_code = SPECIES.ALPHA_3_CODE, PERIOD, VALUE) %>%
  # group by country and species to calculate mean quantity tonnes live weight across years and species
  group_by(UN_Code, species_code) %>%
  summarize(tonnes_lw = mean(VALUE, na.rm=TRUE)) %>%
  ungroup()

# read in country code csv
country <- read_csv(paste0(fao_data_path, "CL_FI_COUNTRY_GROUPS.csv")) %>%
  dplyr::select(UN_Code, UN_country_name = Name_En)

# check for duplicates
sum(duplicated(country$UN_Code))

# read in species information for fish
species <- read_csv(paste0(fao_data_path, "CL_FI_SPECIES_GROUPS.csv")) %>%
  # select species code, taxa code, taxa name, major group. ISSCAAP Group, and scientific name
  dplyr::select(species_code = `3A_Code`, Taxonomic_Code, taxa_name = Name_En, Major_Group_En, ISSCAAP_Group_En, Scientific_Name) %>%
  unique

# check for duplicates
sum(duplicated(species$species_code))

# explore differences between species code in production and species data
setdiff(production_clean$species_code, species$species_code)
# explore difference between UN code from production bs UN code from countries df
setdiff(production_clean$UN_Code, country$UN_Code)

# join production data with country data for names of countries to be added
production <- left_join(production_clean, country, by="UN_Code") %>%
  # join with the species code data to get species information
  left_join(species, by="species_code") %>%
  # fix Turkey name for iso3c
  mutate(UN_country_name = ifelse(UN_country_name=="Türkiye", "Turkey", UN_country_name)) %>%
  # get iso3c from country name
  mutate(iso3c = countrycode(UN_country_name, origin="country.name", destination = "iso3c"),
                       country_name = countrycode(iso3c, origin= "iso3c", destination = "country.name")) %>%
  # these are older country designations with no production for years of interest
  filter(!is.na(iso3c)) %>% 
  filter(tonnes_lw > 0)

```


## Assigning taxonomic categories

NOTE: Considered including the "Sturgeons, paddlefishes" as "Tilapias and other cichlids", but didn't due to Jessica's comments:
- I suspect a lot of that production is for caviar and the pressures for tilapia probably doesn't apply well.
- It also looks like there is quite a bit of production of juveniles for stocking recreational fisheries.
```{r}
# read in groupings data
groups <- read_csv(paste0(fw_data_path, "taxa_isscaap_groupings.csv")) %>%
  filter(environment=="freshwater")
unique(groups$isscaap_group)

# see what groupings match in our data
intersect(production$ISSCAAP_Group_En, groups$isscaap_group)
# see what groups are included in production data but not grouping data
in_fao <- setdiff(production$ISSCAAP_Group_En, groups$isscaap_group)
# see what groups are included in grouping data but not production data
setdiff(groups$isscaap_group, production$ISSCAAP_Group_En)

# review what groups we are missing from grouping data
tmp <- filter(production, ISSCAAP_Group_En %in% in_fao)
unique(tmp$ISSCAAP_Group_En)

# explore pearls data
filter(production, ISSCAAP_Group_En %in% "Pearls, mother-of-pearl, shells" ) %>% data.frame()

## change some groupings here
production_update <- production %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Shrimps, prawns", "Freshwater crustaceans", ISSCAAP_Group_En)) %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Miscellaneous coastal fishes" & taxa_name =="Obscure pufferfish", 
                                   "Miscellaneous freshwater fishes", 
                                   ISSCAAP_Group_En)) %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Miscellaneous coastal fishes" & taxa_name =="Flathead grey mullet", 
                                   "Miscellaneous freshwater fishes", 
                                   ISSCAAP_Group_En)) %>%
  mutate(ISSCAAP_Group_En = ifelse(ISSCAAP_Group_En=="Miscellaneous coastal fishes" & taxa_name =="Mullets nei", 
                                   "Miscellaneous freshwater fishes", 
                                   ISSCAAP_Group_En)) 

# see what groupings match in our data
intersect(production_update$ISSCAAP_Group_En, groups$isscaap_group)
# see what groups are included in production data but not grouping data
in_fao <- setdiff(production_update$ISSCAAP_Group_En, groups$isscaap_group)
# see what groups are included in grouping data but not production data
setdiff(groups$isscaap_group, production$ISSCAAP_Group_En)


## now make some adjustments to match Jessica's groups
hypoph_carp <- unique(grep("Hypophthalmichthys", production_update$Scientific_Name, value=TRUE))

production_update <- production_update %>%
  #make stressor groups that match our pressures data
  mutate(stressor_group = ifelse(taxa_name %in% c("Bighead carp", "Silver carp", "Silver, bighead carps nei"), 
                                 "hypoph_carp", 
                                 NA)) %>%
  mutate(stressor_group = ifelse(Scientific_Name %in% hypoph_carp, 
                                 "hypoph_carp", 
                                 stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Carps, barbels and other cyprinids" & is.na(stressor_group), 
                                 "oth_carp", 
                                 stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Tilapias and other cichlids" & is.na(stressor_group), 
                                 "tilapia", 
                                 stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Miscellaneous freshwater fishes" & is.na(stressor_group), 
                                 # most are catfish per Jessica
                                 "catfish", 
                                 stressor_group)) %>%  
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Salmons, trouts, smelts" & is.na(stressor_group), 
                                 "trout", 
                                 stressor_group)) %>%
  mutate(stressor_group = ifelse(taxa_name == "Milkfish" & is.na(stressor_group), 
                                 "milkfish", 
                                 stressor_group)) %>%
  mutate(stressor_group = ifelse(ISSCAAP_Group_En == "Miscellaneous diadromous fishes" & is.na(stressor_group), 
                                 "misc_diad", 
                                 stressor_group))

# see how much of the tonnes_lw data is accounted for in our groupings
sum(production_update$tonnes_lw[!is.na(production_update$stressor_group)])/sum(production_update$tonnes_lw)
## accounts for about 91% of production

## summarize by Jessica's groupings
tonnes_lw <- production_update %>%
  # remove NA stressor_group rows
  filter(!is.na(stressor_group)) %>%
  # for each country and stressor group
  group_by(iso3c, country_name, stressor_group) %>%
  # calculate the total tonnes_lw produced
  summarize(tonnes_lw = sum(tonnes_lw)) %>%
  arrange(-tonnes_lw)

```

## Calculate Feed Composition for FW Aquaculture

### Crop feed composition
- Estimate proportion of oilcrops vs. grains
Mostly from food_systems/feed/diet_composition_wrangling.xlsx
For catfish: mostly: https://agric4profits.com/the-ideal-feeds-for-catfish/
https://thefishsite.com/articles/catfish-nutrition-feeds-and-feed-formulation
https://www.fao.org/3/Q3567E/q3567e05.htm
rice bran: https://onlinelibrary.wiley.com/doi/10.1111/J.1365-2109.2011.03048.X
rice bran: https://fishfeedmachinery.com/Solution/nutritional-fish-feed-formulation.html

For tilapia: https://www.fao.org/fishery/affris/species-profiles/nile-tilapia/feed-formulation/en/

- Ignores animal products (beyond fish oil fish meal)

- Use fofm data from Halley for each group and then distribute based on tonnes of production in each country

- rescale hypoph carps so feed sums to 1

FCR stands for feed conversion ratio and is the amount of crop (ex. kg or tonne) required to produce 1 unit of live weight of fish (in either kg or tonne). It is essentially a unitless ratio, however the fish component needs to be in live weight equivalent.
```{r}
# taxa list of our groupings
taxa_list <- c("oth_carp", "hypoph_carp", "catfish", "tilapia", "trout", "milkfish", "misc_diad")

crop_cats <- read_csv(paste0(fw_data_path, "crop_breakout.csv"))
# list.files(paste0(pressure_data_path, "fresh_water_aqua_pressures_Jessica/")

feed <- read_csv(paste0(pressures_data_path, "fresh_water_aqua_pressures_Jessica/aquaculture_group_fcr_feed_averages.csv")) %>%
  # filter for the fw taxa we are including
  filter(taxa %in% taxa_list) %>%
  rowwise() %>%
  # calculate the total diet allocation for each taxa group
  mutate(total = feed_soy + feed_crops + feed_fmfo + feed_animal) %>%
  # calculate the percentage makeup of total feed for each category
  mutate(feed_soy = feed_soy/total,
         feed_crops = feed_crops/total,
         feed_fmfo = feed_fmfo/total,
         feed_animal = feed_animal/total) %>%
  # join with crops data
  left_join(crop_cats, by="taxa") %>%
  # using the feed_crops category, calculate the makeup of the crops group by oils and grains
  mutate(feed_grain = feed_crops * grains) %>%
  mutate(feed_oil = feed_crops * oil_crops) %>%
  # recalculate the the total diet allocation with the new feed_grain and feed_oil categories
  mutate(total = feed_soy + feed_grain + feed_oil + feed_fmfo + feed_animal) %>%
  # select required columns
  dplyr::select(stressor_group = taxa, fcr, feed_soy,feed_grain, feed_oil, feed_fmfo, total)

```

### Crop feed quantity
Calculate tonnes of each crop feed category for each country and taxa group based on fcr, diet composition, and tonnes of production.
```{r}
# join the tonnes_lw data with the feed data
feed_tonnes_pre <- left_join(tonnes_lw, feed, by="stressor_group") %>%
  rowwise() %>%
  # for each row calculate the tonnes of soy, grain, and oil feed required
  mutate(tonnes_feed_soy = tonnes_lw * fcr * feed_soy) %>%
  mutate(tonnes_feed_grain = tonnes_lw * fcr * feed_grain) %>%
  mutate(tonnes_feed_oil = tonnes_lw * fcr * feed_oil) %>%
  select(iso3c, country_name, stressor_group, tonnes_feed_soy, tonnes_feed_grain, tonnes_feed_oil) %>%
  pivot_longer(c("tonnes_feed_soy", "tonnes_feed_grain", "tonnes_feed_oil"), names_to = "feed_type", values_to = "tonnes_feed")

```

Correct for averaged loss during processing. This is based on averaged data for each feed category in food_systems/feed/data/feed_extraction_rates.csv. This is a rough estimate but is the best data available.
```{r}
# make a data frame of percent of crop lost during processing
crop_loss <- data.frame(feed_type = c("tonnes_feed_soy", "tonnes_feed_grain", "tonnes_feed_oil"), 
                        loss_percent = c(3, 2, 9))

# join crop loss with calculated tonnes of feed type to back calculate total crop tonnes required
crop_feed_tonnes <- left_join(feed_tonnes_pre, crop_loss, by="feed_type") %>%
  # adjust tonnes_product needed by using loss percentage to increase total by crop lost during processing
  mutate(tonnes_product = tonnes_feed * 100/(100 - loss_percent)) %>%
  mutate(feed_category = ifelse(feed_type == "tonnes_feed_soy", "soy", NA)) %>%
  mutate(feed_category = ifelse(feed_type == "tonnes_feed_grain", "grain", feed_category)) %>%
  mutate(feed_category = ifelse(feed_type == "tonnes_feed_oil", "oil", feed_category)) %>%
  select(iso3c, country_name, taxa=stressor_group, feed_category, tonnes=tonnes_product)

write_csv(crop_feed_tonnes, paste0(save_path, "fw_aqua_crop_feed.csv"))
```

## Fish oil fish meal consumption - fresh water aquaculture

### Read in FOFM consumption for marine aquaculture
We will use the values from Halley et al. (https://knb.ecoinformatics.org/view/doi:10.5063/F1T43R7N) as FOFM consumption for species and distribute consumption of fofm based on production data from above. The fofm consumption data is in live weight equivalent tonnes, this could be converted into true FOFM tonnes with a 28% reduction in processing loss (reference food_pressures paper). However, the FOFM trade data is in live weight equivalents so for now we will keep in true_tonnes_fish for our calculations.
```{r}
# fofm consumption reference
fofm_reference <- read_csv(paste0(fw_data_path, "msleckman.45.1_fofm_consumption.csv")) %>%
  filter(!is.na(fw_aquaculture)) %>% 
  # current mean represents the live weight of fish for fofm consumption
  # rename current mean to live weight of fish 
  rename(lw_tonnes_fish = Current_Mean)   

```

### Adjust Consumption
Need to make a few adjustments to the FOFM consumption data to calculate values for all of the species we have trade/pressure data for

1. We use catfish pressures for ISSCAAP "Miscellaneous freshwater fishes". These are mostly catfish, and difficult to identify but there is reported aquaculture for: "Miscellaneous coastal fishes" and "Marine fishes not identified", but these end up  being <<1% of total production so we can ignore these contributions. This data check is done below

```{r}
# filter production data to get "Miscellaneous coastal fishes" and "Marine fishes not identified"
fw_fishes_not_inc <- filter(production_update, is.na(stressor_group)) %>%
  filter(ISSCAAP_Group_En %in% c("Miscellaneous coastal fishes", "Marine fishes not identified"))

fw_fishes_inc <- filter(production_update, ISSCAAP_Group_En %in% c("Miscellaneous coastal fishes", "Marine fishes not identified", "Miscellaneous freshwater fishes"))

sum(fw_fishes_not_inc$tonnes_lw)/sum(fw_fishes_inc$tonnes_lw) # <<1% can assume all fofm goes to catfish
```

2. fofm reference includes diadromous fishes, which the pressures break into two categories: milkfish and misc_diads (Barramundi and a very small amount of striped bass). Data check is done below

```{r}
# view  makeup diadromous fish groups
tmp <- filter(production_update, ISSCAAP_Group_En %in% "Miscellaneous diadromous fishes") %>% data.frame()

# make separate for milkfish
milkfish <- filter(production_update, stressor_group == "milkfish")
# make separate for mis_diad
other_diad <- filter(production_update, stressor_group == "misc_diad")

# how much of diadromous fishes is milkfish by live wieght
sum(milkfish$tonnes_lw)/(sum(milkfish$tonnes_lw)+sum(other_diad$tonnes_lw)) # 90% milkfish!

# but need to take into account proportion fofm and fcr in diet from Jessica's data:
milkfish_prop <- (sum(milkfish$tonnes_lw)*feed$feed_fmfo[feed$stressor_group == "milkfish"]*feed$fcr[feed$stressor_group == "milkfish"])/ ((sum(milkfish$tonnes_lw)*feed$feed_fmfo[feed$stressor_group == "milkfish"]*feed$fcr[feed$stressor_group == "milkfish"]) + (sum(other_diad$tonnes_lw) * feed$feed_fmfo[feed$stressor_group == "misc_diad"]*feed$fcr[feed$stressor_group == "misc_diad"])) # 75% fofm by milkfish!
```

3. Carps were divided into hypoph_carp and other carp

```{r}
# separate carp taxa
hypoph_carp <- filter(production_update, stressor_group == "hypoph_carp")
oth_carp <- filter(production_update, stressor_group == "oth_carp")
# how much of carp is hypoph_carp
sum(hypoph_carp$tonnes_lw)/(sum(hypoph_carp$tonnes_lw)+sum(oth_carp$tonnes_lw)) # 28% hypoph_carp

# adjust amount of fofm using pcr and diet quantity
hypoph_carp_prop <- (sum(hypoph_carp$tonnes_lw) * feed$feed_fmfo[feed$stressor_group == "hypoph_carp"] * feed$fcr[feed$stressor_group == "hypoph_carp"])/ (sum(hypoph_carp$tonnes_lw)*feed$feed_fmfo[feed$stressor_group == "hypoph_carp"] * feed$fcr[feed$stressor_group == "hypoph_carp"] + (sum(oth_carp$tonnes_lw)*feed$feed_fmfo[feed$stressor_group == "oth_carp"] * feed$fcr[feed$stressor_group == "oth_carp"])) # 4% fofm to hypoph carp
```


Using the above steps, divide up the fofm consumption data based on the species makeup from our data.
```{r}
divided_fofm <- data.frame(fw_aquaculture=c("milkfish", "misc_diad", "hypoph_carp", "oth_carp"),
                           prop =c(milkfish_prop, (1-milkfish_prop), hypoph_carp_prop, (1 - hypoph_carp_prop)))

fofm <- left_join(fofm_reference, divided_fofm, by="fw_aquaculture") %>%
  mutate(prop = ifelse(is.na(prop), 1, prop)) %>%
  # use true tonnes fish because trade data is in live weight tonnes
  mutate(taxa_lw_tonnes_fofm_consumption = lw_tonnes_fish*prop) %>%
  select(stressor_group = fw_aquaculture, taxa_lw_tonnes_fofm_consumption)

```

Distribute fofm consumption based on production by country

```{r}
# using the tonnes_lw production data from above
fofm_tonnes <- tonnes_lw %>%
  group_by(stressor_group) %>%
  # calculate the total tonnes_lw produced for each stressor_group
  mutate(global_tonnes=sum(tonnes_lw)) %>%
  ungroup() %>%
  rowwise() %>%
  # calculate the proportion of that stressor_group that was produced by that country
  mutate(prop_prod = tonnes_lw/global_tonnes) %>%
  # join with tonnes of fofm consumed by each stressor group
  left_join(fofm, by="stressor_group") %>%
  # calculate the tonnes_fofm consumed by that country for FW aquaculture
  mutate(lw_tonnes_fofm = taxa_lw_tonnes_fofm_consumption*prop_prod) %>%
  select(iso3c, country_name, taxa=stressor_group, lw_tonnes_fofm)  

write_csv(fofm_tonnes, paste0(save_path, "fw_fofm_consumption.csv")) 
```