---
title: "Calculate the proportion of crop and fofm used for each animal for feed in a country"
date: "2023-10-03"
output: html_document
---

Objective: Calculate the proportion of crop and fish oil fish meal (FOFM) a country consumes that goes to each animal for feed in a country. 
## Setup

Load libraries
```{r}
library(tidyverse)
```

Data Paths
```{r}
raw_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
animal_consumption_data_path <- paste0(raw_path, "Halpern_feed_consumption/")
fw_aquaculture_consumption_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/03b_fofm_fw_aquaculture_feed/"
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/03c_calculate_prop_crop_and_fofm_used_for_each_animal/"
```

## Crops

### Read in and prepare livestock tonnes of crop consumed per animal per country
Read in tonnes crop consumed per animal group per country data, the data used for this step can be found at the link provided, however it has been saved in the project folder on Aurora: https://github.com/OHI-Science/food_systems/blob/master/feed/data/system_country_mapspam_tonnes_consumption.csv
```{r}
animal_consumption_raw <- read_csv(paste0(animal_consumption_data_path, "system_country_mapspam_tonnes_consumption.csv"))
```

Subset to 5 feed crop categories we're using and aggregate for grains
```{r}
animal_consumption <- animal_consumption_raw %>% 
  filter(SPAM_super == "barl" | SPAM_super == "maiz" | SPAM_super == "ocer" | SPAM_super == "rice" |
           SPAM_super == "sorg" | SPAM_super == "whea" | SPAM_super == "xmil" | SPAM_super == "soyb" |
           SPAM_super == "xoil" | SPAM_super == "oilp" | SPAM_super == "xpul") %>% 
  mutate(feed_category = case_when(SPAM_super == "barl" ~ "grain",
                                   SPAM_super == "maiz" ~ "grain",
                                   SPAM_super == "ocer" ~ "grain",
                                   SPAM_super == "rice" ~ "grain",
                                   SPAM_super == "sorg" ~ "grain",
                                   SPAM_super == "whea" ~ "grain",
                                   SPAM_super == "xmil" ~ "grain",
                                   SPAM_super == "soyb" ~ "soybean",
                                   SPAM_super == "xoil" ~ "oilcrop",
                                   SPAM_super == "oilp" ~ "palmoil",
                                   SPAM_super == "xpul" ~ "pulse",
                                   TRUE ~ as.character(NA)))
```


Match animal categories to what we're using with FAO trade data and the pressures data.

Categories from trade:

Livestock:
chickens_eggs
chickens_meat
cows_meat
goats_meat
cows_milk
pigs_meat
sheep_meat
sheep_milk
buffaloes_milk

Mariculture (we have pressures data for bivalves but no feed data, because they don't need feed because they filter):
Renamed animal_agg is on the right side of the equations
crustaceans_aquaculture_meat = crustaceans_aquaculture
marine-fish-general_aquaculture_meat = marine_fish_general_aquaculture
salmon_aquaculture_meat = salmonids_aquaculture
shrimp_aquaculture_meat = shrimps_prawns_aquaculture
tuna_aquaculture_meat = tuna_aquaculture

```{r}
# how many rows are 0 tonnes crop product consumed by that animal group?
length(which(animal_consumption$tonnes_product == 0))
# 22648 rows have 0 tonnes (63% of data)

#table(animal_consumption$animal_system)

# Match animal categories to what we're using with FAO data
animal_consumption_agg <- animal_consumption %>% 
  mutate(animal_agg = case_when(str_detect(animal_system, regex("(?=.*cows)(?=.*meat)", ignore_case = TRUE)) ~ "cows_meat",
                                str_detect(animal_system, regex("(?=.*cows)(?=.*milk)", ignore_case = TRUE)) ~ "cows_milk",
                                str_detect(animal_system, regex("(?=.*goat)(?=.*meat)", ignore_case = TRUE)) ~ "goats_meat",
                                str_detect(animal_system, regex("(?=.*buffaloes)(?=.*milk)", ignore_case = TRUE)) ~ "buffaloes_milk",
                                str_detect(animal_system, regex("(?=.*sheep)(?=.*meat)", ignore_case = TRUE)) ~ "sheep_meat",
                                str_detect(animal_system, regex("(?=.*sheep)(?=.*milk)", ignore_case = TRUE)) ~ "sheep_milk",
                                str_detect(animal_system, regex("(?=.*pig)(?=.*meat)", ignore_case = TRUE)) ~ "pigs_meat",
                                animal_system == "chickens_industrial_eggs" ~ "chickens_eggs",
                                animal_system == "chickens_industrial_meat" ~ "chickens_meat",
                                animal_system == "crustaceans_aquaculture_meat" ~ "crustaceans_mariculture",
                                animal_system == "marine-fish-general_aquaculture_meat" ~ "marine_fish_general_mariculture",
                                animal_system == "salmon_aquaculture_meat" ~ "salmonids_mariculture",
                                animal_system == "shrimp_aquaculture_meat" ~ "shrimps_prawns_mariculture",
                                animal_system == "tuna_aquaculture_meat" ~ "tuna_mariculture",
                                TRUE ~ animal_system))

#table(animal_consumption_agg$animal_agg)

# remove all buffaloes meat, goats milk, and chickens_backyard_eggs&meat as we do not have trade data or pressures data for these
animal_consumption_agg <- animal_consumption_agg %>% 
  filter(!animal_agg == "buffaloes_grassland_meat") %>% 
  filter(!animal_agg == "buffaloes_mixed_meat") %>% 
  filter(!animal_agg == "chickens_backyard_eggs&meat") %>% 
  filter(!animal_agg == "goats_grassland_milk") %>% 
  filter(!animal_agg == "goats_mixed_milk") 

#table(animal_consumption_agg$animal_agg)
```

### Read in and prepare fw aquaculture tonnes of crop consumed per animal per country
Read in inland aquaculture feed consumption data produced in script 03b.
```{r}
fw_aqua_consumption <- read_csv(paste0(fw_aquaculture_consumption_path, "fw_aqua_crop_feed.csv"))
```

Convert to inland aquaculture trade categories

Key for joining trade and consumption - updated data (left is trade matrix categories, right is pressures categories):
oth_carp = oth_carp 
hypoph_carp = hypoph_carp
miscellaneous diadromous fishes = misc_diad + milkfish
salmons, trouts, smelts = trout
miscellaneous freshwater fishes = catfish
tilapias and other cichlids = tilapia

We are adding "fw_aquaculture" to the beginning of the items for freshwater aquaculture to make them easily distinguishable from marine fisheries and aquaculture.
```{r}
fw_aqua_consumption_relabel <- fw_aqua_consumption %>% 
  mutate(nceas_group = case_when(taxa == "catfish" ~ "fw_aquaculture_miscellaneous_freshwater_fishes",
                                 taxa == "hypoph_carp" ~ "fw_aquaculture_hypoph_carp",
                                 taxa == "milkfish" ~ "fw_aquaculture_miscellaneous_diadromous_fishes",
                                 taxa == "misc_diad" ~ "fw_aquaculture_miscellaneous_diadromous_fishes",
                                 taxa == "oth_carp" ~ "fw_aquaculture_oth_carp",
                                 taxa == "tilapia" ~ "fw_aquaculture_tilapias_and_other_cichlids",
                                 taxa == "trout" ~ "fw_aquaculture_salmons_trouts_smelts",
                                 TRUE ~ as.character(NA)))
#summary(fw_aqua_consumption_relabel)
#table(fw_aqua_consumption_relabel$nceas_group)
```

### Join inland aquaculture, marine fisheries and aquaculture, and livestock crop feed quantities.
Bind inland aquaculture feed with livestock/mariculture/wild fisheries feed
```{r}
animal_consumption_category_totals <- animal_consumption_agg %>% 
  # calculate tonnes feed consumed per crop, animal system, and country
  group_by(iso3c, feed_category, animal_agg) %>% 
  summarise(tonnes_feed = sum(tonnes_product)) %>% 
  rename(animal_group = animal_agg) %>% 
  ungroup()

fw_aqua_consumption_category_totals <- fw_aqua_consumption_relabel %>% 
  # calculate tonnes feed consumed per crop, animal system (aquaculture), and country
  group_by(iso3c, feed_category, animal_group = nceas_group) %>% 
  summarise(tonnes_feed = sum(tonnes)) %>% 
  ungroup()

# combine feed dfs (livestock/fisheries/mariculture and fw aquaculture)
tonnes_crop_feed_consumed <- rbind(animal_consumption_category_totals, fw_aqua_consumption_category_totals)

# check data
# tonnes_crop_feed_consumed %>%
#   group_by(animal_group) %>%
#   summarise(sum = sum(tonnes_feed)) %>%
#   left_join(animal_consumption_agg %>%
#               group_by(animal_agg) %>%
#               summarise(livestock_sum = sum(tonnes_product)),
#             by = c("animal_group" = "animal_agg")) %>%
#   left_join(fw_aqua_consumption_relabel %>%
#               group_by(nceas_group) %>%
#               summarise(fw_sum = sum(tonnes)),
#             by = c("animal_group" = "nceas_group")) %>%
#   rowwise() %>%
#   mutate(change = sum - sum(livestock_sum, fw_sum, na.rm = TRUE))
# no differences in total tonnage after joining!
```

### Calculate proportion of each crop that goes to each animal system within a country

```{r}

# rows where total crop = 0 means this country doesn't use any of this crop as feed
crop_feed_animal_proportions <- tonnes_crop_feed_consumed %>% 
  group_by(iso3c, feed_category) %>% 
  #calculate the total tonnes of crop a country consumes for feed per this data
  mutate(total_crop = sum(tonnes_feed)) %>% 
  ungroup() %>% 
  # calculate the proportion of each feed crop that goes to each animal
  mutate(prop_crop_feed_per_animal = case_when(tonnes_feed != 0 ~ tonnes_feed/total_crop,
                                               TRUE ~ 0)) %>% 
  # removed tonnage from data source as we were using that to calculate the proportions
  select(-tonnes_feed, -total_crop)

# check - should all sum to 1 or 0 (country doesn't use any of this crop as feed for any animals)
check_props <- crop_feed_animal_proportions %>% 
  group_by(iso3c, feed_category) %>% 
  summarise(sum = sum(prop_crop_feed_per_animal)) %>% 
  ungroup()
range(check_props$sum)

# pivot crop_feed_animal_proportions to wide for joining
crop_feed_animal_proportions_wide <- crop_feed_animal_proportions %>% 
  pivot_wider(names_from = animal_group,
              values_from = prop_crop_feed_per_animal)

# Pivoting wider produced a lot of NA columns because there no data for some animals and feed_categories
crop_feed_animal_proportions_wide[is.na(crop_feed_animal_proportions_wide)] <- 0

```

### Save
```{r}
write_csv(x = crop_feed_animal_proportions_wide,
          file = paste0(save_path, "crop_feed_animal_proportions_wide.csv"))
```

## FOFM
This section will largely do the same as above does for crops except for Fish Oil and Fish Meal. In step 03b we calculated to quantity of FOFM required for each country's production of freshwater aquaculture. We will combine this is the tonnage of FOFM used in other feed streams and calculate each country's proportion of FOFM that goes to each animal system.

### Read in and prepare livestock and marine aquaculture tonnes of FOFM consumed per animal per country
The FOFM consumption data (`FMFO_country_data.csv`) is downloaded from: https://github.com/OHI-Science/food_systems/blob/master/feed/data/FMFO_country_data.csv 

The tonnes column represents the tonnes of live weight consumed of the Fish that make up FOFM.
```{r}
FOFM_consumption_raw <- read_csv(paste0(animal_consumption_data_path, "FMFO_country_data.csv"))

# the data for FOFM consumption for marine salmon aquaculture includes
# freshwater aquaculture propduction tonnes. Below we will reduce all 
# salmon consumption values by 25% according to Mel (chicken/salmon paper)
FOFM_consumption <- FOFM_consumption_raw %>% 
  mutate(tonnes = if_else(condition = system == "salmon_aquaculture_meat",
                          # 25% reduction
                          true = tonnes*.75,
                          # keep tonnes the same for all other species
                          false = tonnes))

# check out species included
unique(FOFM_consumption$system)

```

Match animal categories to what we're using with FAO trade data and the pressures data. The left side of the equations is how the data is read in and the right side of the equations is the renamed animal_agg group to match the trade and pressures data

Categories from trade included in FOFM feed:

Livestock:
chickens_industrial_eggs = chickens_eggs
chickens_industrial_meat = chickens_meat
pigs_intermediate_meat + pigs_backyard_meat + pigs_industrial_meat = pigs_meat

Mariculture:
crustaceans_aquaculture_meat = crustaceans_aquaculture
marine-fish-general_aquaculture_meat = marine_fish_general_aquaculture
salmon_aquaculture_meat = salmonids_aquaculture
shrimp_aquaculture_meat = shrimps_prawns_aquaculture
tuna_aquaculture_meat = tuna_aquaculture

```{r}
# Match animal categories to what we're using with FAO data
FOFM_consumption_aggregate <- FOFM_consumption %>% 
  mutate(animal_aggregate = case_when(str_detect(system, regex("(?=.*pig)(?=.*meat)", ignore_case = TRUE)) ~ "pigs_meat",
                                      system == "chickens_industrial_eggs" ~ "chickens_eggs",
                                      system == "chickens_industrial_meat" ~ "chickens_meat",
                                      system == "crustaceans_aquaculture_meat" ~ "crustaceans_mariculture",
                                      system == "marine-fish-general_aquaculture_meat" ~ "marine_fish_general_mariculture",
                                      system == "salmon_aquaculture_meat" ~ "salmonids_mariculture",
                                      system == "shrimp_aquaculture_meat" ~ "shrimps_prawns_mariculture",
                                      system == "tuna_aquaculture_meat" ~ "tuna_mariculture",
                                      TRUE ~ system))

```

### Read in and prepare fw aquaculture tonnes of FOFM consumed per animal per country
Read in inland aquaculture feed consumption data produced in script 03b.
```{r}
fw_aqua_FOFM_consumption <- read_csv(paste0(fw_aquaculture_consumption_path, "fw_fofm_consumption.csv"))

# check out species included
unique(fw_aqua_FOFM_consumption$taxa)
```

Match animal categories to what we're using with FAO trade data and the pressures data. The left side of the equations is how the data is read in and the right side of the equations is the renamed animal_agg group to match the trade and pressures data.

FW Aquaculture:
oth_carp = oth_carp
hypoph_carp = hypoph_carp
catfish = miscellaneous_freshwater_fishes
tilapia = tilapias_and_other_cichlids
trout = salmons_trouts_smelts
milkfish + misc_diad = miscellaneous_diadromous_fishes

We are adding "fw_aquaculture" to the beginning of the items for freshwater aquaculture to make them easily distinguishable from marine fisheries and aquaculture.
```{r}
fw_aqua_FOFM_consumption_relabel <- fw_aqua_FOFM_consumption %>% 
  mutate(animal_aggregate = case_when(taxa == "catfish" ~ "fw_aquaculture_miscellaneous_freshwater_fishes",
                                      taxa == "hypoph_carp" ~ "fw_aquaculture_hypoph_carp",
                                      taxa == "milkfish" ~ "fw_aquaculture_miscellaneous_diadromous_fishes",
                                      taxa == "misc_diad" ~ "fw_aquaculture_miscellaneous_diadromous_fishes",
                                      taxa == "oth_carp" ~ "fw_aquaculture_oth_carp",
                                      taxa == "tilapia" ~ "fw_aquaculture_tilapias_and_other_cichlids",
                                      taxa == "trout" ~ "fw_aquaculture_salmons_trouts_smelts",
                                      TRUE ~ as.character(NA)))

```

### Join inland aquaculture, marine fisheries and aquaculture, and livestock crop feed quantities.
Bind inland aquaculture feed with livestock/mariculture/wild fisheries feed
```{r}
FOFM_consumption_totals <- FOFM_consumption_aggregate %>% 
  # calculate tonnes fofm feed consumed by, animal system, and country
  group_by(iso3c, animal_group = animal_aggregate) %>% 
  summarise(lw_tonnes_fofm = sum(tonnes)) %>% 
  ungroup()

fw_aqua_FOFM_consumption_totals <- fw_aqua_FOFM_consumption_relabel %>% 
  # calculate tonnes fofm feed consumed by, animal system, and country
  group_by(iso3c, animal_group = animal_aggregate) %>% 
  summarise(lw_tonnes_fofm = sum(lw_tonnes_fofm)) %>% 
  ungroup()

# combine feed dfs (livestock/fisheries/mariculture and fw aquaculture)
lw_tonnes_fofm_feed_consumed <- rbind(FOFM_consumption_totals, fw_aqua_FOFM_consumption_totals)

# check data
# lw_tonnes_fofm_feed_consumed %>%
#   group_by(animal_group) %>%
#   summarise(sum = sum(lw_tonnes_fofm)) %>%
#   left_join(FOFM_consumption_aggregate %>%
#               group_by(animal_aggregate) %>%
#               summarise(livestock_sum = sum(tonnes)),
#             by = c("animal_group" = "animal_aggregate")) %>%
#   left_join(fw_aqua_FOFM_consumption_relabel %>%
#               group_by(animal_aggregate) %>%
#               summarise(fw_sum = sum(lw_tonnes_fofm)),
#             by = c("animal_group" = "animal_aggregate")) %>%
#   rowwise() %>%
#   mutate(change = sum - sum(livestock_sum, fw_sum, na.rm = TRUE))
# no differences in total tonnage after joining!

# calculate the total lw tonnes consumed by each country
# this will be used to allocate the correct amount of tonnage 
# FOFM traded to a country towards feed processes
lw_tonnes_fofm_feed_consumed_per_country <- lw_tonnes_fofm_feed_consumed %>% 
  group_by(iso3c) %>% 
  #calculate the total tonnes of crop a country consumes for feed per this data
  summarise(total_lw_tonnes_fofm = sum(lw_tonnes_fofm)) %>% 
  ungroup()

# save total tonnage
write_csv(x = lw_tonnes_fofm_feed_consumed_per_country,
          file = paste0(save_path, "lw_tonnes_fofm_feed_consumed_per_country.csv"))
```

### Calculate proportion of each crop that goes to each animal system within a country

```{r}

# rows where total lw_tonnes_fofm = 0 means this country doesn't use any fofm as feed
fofm_feed_animal_proportions <- lw_tonnes_fofm_feed_consumed %>% 
  group_by(iso3c) %>% 
  #calculate the total tonnes of crop a country consumes for feed per this data
  mutate(total_lw_fofm = sum(lw_tonnes_fofm)) %>% 
  ungroup() %>% 
  # calculate the proportion of each feed crop that goes to each animal
  mutate(prop_fofm_feed_per_animal = case_when(lw_tonnes_fofm != 0 ~ lw_tonnes_fofm/total_lw_fofm,
                                               TRUE ~ 0)) %>% 
  # removed tonnage from data source as we were using that to calculate the proportions
  select(-lw_tonnes_fofm, -total_lw_fofm)

# check - should all sum to 1 or 0 (country doesn't use any fofm feed for any animals)
check_props <- fofm_feed_animal_proportions %>% 
  group_by(iso3c) %>% 
  summarise(sum = sum(prop_fofm_feed_per_animal)) %>% 
  ungroup()
range(check_props$sum)

# pivot crop_feed_animal_proportions to wide for joining
fofm_feed_animal_proportions_wide <- fofm_feed_animal_proportions %>% 
  pivot_wider(names_from = animal_group,
              values_from = prop_fofm_feed_per_animal) %>% 
  mutate(feed_category = "fofm",
         .after = iso3c)

# Pivoting wider produced a lot of NA columns because there no data for some animals and feed_categories
fofm_feed_animal_proportions_wide[is.na(fofm_feed_animal_proportions_wide)] <- 0

```

### Save
```{r}
write_csv(x = fofm_feed_animal_proportions_wide,
          file = paste0(save_path, "fofm_feed_animal_proportions_wide.csv"))
```

