---
title: "01a-format-fao-trade"
date: "2023-05-09"
editor_options: 
  chunk_output_type: console
---

Objective: Format the 2017 FAO production, commodity balances, and detailed trade matrix datasets to match Kastner's data that is input into the trade matrix script.

Update: Formatted data for 2015, 2016, 2018, and 2019 on Nov 28, 2022.

```{r}
library(tidyverse)
library(data.table)
```


### Setup
```{r}
raw_folder_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
kastner_data_path <- paste0(raw_folder_path, "kastner_2022/")
fao_data_path <- paste0(raw_folder_path, "FAO/2017_FAO_data/")
production_data_path <- paste0(raw_folder_path, "FAO/production_data/")
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01a_format_fao_trade/"
```

### Detailed trade matrix

Trade data - get error on FAO website when trying to download just one year of data. We have to use the bulk download and download all years of data and then later subset to 2017.

We need to gapfill missing trade data with trading partners (FAO does this for some, but not all).


Kastner Trade Matrix example to understand how to set up the other data
```{r}
#read in Kastner detailed trade matrix data (from FAO) - This is to understand what we want the output data to look like
k_trade_matrix <- fread(paste0(kastner_data_path, "Trade_DetailedTradeMatrix_E_All_Data_(Normalized).csv"))
head_k_trade_matrix <- head(k_trade_matrix)
# k_trade_matrix is not used anywhere else
rm(k_trade_matrix)
```

Trade data - Gapfill imports with export data
```{r}
#read in recently downloaded FAO detailed trade matrix data
fao_trade_matrix_raw <- fread(paste0(fao_data_path, "Trade_DetailedTradeMatrix_E_All_Data_(Normalized)_7_14_22.csv"))

fao_trade_quantity_gapfill <- function(fao_trade_matrix_raw, year) {
  # subset columns to match Kastner data columns
  trade_matrix <- fao_trade_matrix_raw %>% 
  select(-c(`Reporter Country Code (M49)`, `Partner Country Code (M49)`, `Item Code (CPC)`)) %>% 
  filter(Year == year)
  
  # want to prioritize import data and only gapfill with export data
  # from Kastner paper:
  # "There are, for example, sometimes differences between a reported import from country A importing from country B, and the corresponding report of an export from country B to country A. There are also sometimes trade flows missing from reports (Gehlhar 1996). To harmonize these reports, we decided to prioritize import reports as being more reliable and only used reported exports to fill missing flows. This is a reasonable assumption and feasible for such a huge dataset."
  
  # separate import and export data into separate dataframes
  # rename export data so it can be merged with import data, make the columns match up and switch the reporter and partner country names
  export <- trade_matrix %>% 
    filter(`Element Code` == 5910) %>% 
    rename("Partner Countries_1" = "Reporter Countries") %>%
    rename("Partner Country Code_1" = "Reporter Country Code") %>% 
    rename("Reporter Countries" = "Partner Countries") %>% 
    rename("Reporter Country Code" = "Partner Country Code") %>% 
    rename("Partner Countries" = "Partner Countries_1") %>% 
    rename("Partner Country Code" = "Partner Country Code_1") %>% 
    rename("Value_gapfill" = Value) %>% 
    rename("Flag_gapfill" = Flag) %>% 
    select(-c("Element Code", "Element"))
  
  import <- trade_matrix %>% 
    filter(`Element Code` == 5610)

  # join import and export data back together 
  import_export <- full_join(import, export, by = c("Reporter Countries", "Reporter Country Code", "Partner Countries", "Partner Country Code", "Item Code", "Item", "Year Code", "Year", "Unit"))
  
  # fill in missing import quantities with export quantities and make note in Flag column where that's done ("gapfill")
  trade_df_gapfilled <- import_export %>% 
    mutate(Value = ifelse(is.na(Value), Value_gapfill, Value)) %>% 
    mutate(Flag = ifelse(is.na(Flag), "gapfill", Flag)) %>% 
    mutate(`Element Code` = ifelse(is.na(`Element Code`), 5610, `Element Code`)) %>% 
    mutate(Element = ifelse(is.na(Element), "Import Quantity", Element)) %>% 
    select(-c(Value_gapfill, Flag_gapfill))
  
  return(trade_df_gapfilled)
}
```


Gapfill and save as csv for 2015, 2016, 2017, 2018, 2019 data
```{r}
# 2015 data
trade_2015 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2015)

# save df - can take out NAs later
write_csv(trade_2015, paste0(save_path,"Trade_2015_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2016 data
trade_2016 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2016)

# save df - can take out NAs later
write_csv(trade_2016, paste0(save_path,"Trade_2016_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2017 data
trade_2017 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2017)

# save df - can take out NAs later
write_csv(trade_2017, paste0(save_path,"Trade_2017_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2018 data
trade_2018 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2018)

# save df - can take out NAs later
write_csv(trade_2018, paste0(save_path,"Trade_2018_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2019 data
trade_2019 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2019)

# save df - can take out NAs later
write_csv(trade_2019, paste0(save_path,"Trade_2019_DetailedTradeMatrix_Gapfilled_formatted.csv"))
```



Average trade across five years
```{r}
# get the unique values of this row to add back in later
unique(trade_2016$`Element Code`)
unique(trade_2016$Element)
unique(trade_2016$Unit)

# rbind trade df for five years of interest
average_trade <- rbind(trade_2015,
                       trade_2016,
                       trade_2017,
                       trade_2018,
                       trade_2019) %>%
  # group by all columns of specificity except for year and description columns
  group_by(`Reporter Country Code`,
           `Reporter Countries`,
           `Partner Country Code`,
           `Partner Countries`,
           `Item Code`,
           Item) %>% 
  summarise(Value_average = mean(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # add back in three columns that are missing when groupby was done
  # after adding these columns back everything should match the columns of the Kastner matrix
  mutate(`Element Code` = 5610,
         Element = "Import Quantity",
         Unit = "tonnes") 

# Test where the NaN are coming from in the Value_average column
NaN_test <- average_trade %>%  
  filter(is.nan(Value_average))
# NaN only occur in rows where the Value is NA so there is nothing we can really do about this for now
# Also is only for 20,000 rows out of almost two million observations so can probably ignore and filter for these later

write_csv(average_trade, paste0(save_path,"Trade_average_DetailedTradeMatrix_Gapfilled_formatted.csv"))

rm(fao_trade_matrix_raw,
   trade_2015,
   trade_2016,
   trade_2017,
   trade_2018,
   trade_2019,
   NaN_test)
```


### Production data

Data in the bulk download files is shown pivoted by years (wide format), except for the Normalized file which is in database format (long format).

Kastner FAO production data is normalized data (long format) separated into crops, processed crops, and primary livestock. We can download these distinct categories from the FAO website if we do not use the bulk download option. Downloading selected options from the FAO website provides the data in long format (i.e. normalized).

FAO products included in primary crops, processed crops, primary livestock, and processed livestock.
Crop and livestock statistics are recorded for 278 products, covering the following categories: 
1) CROPS PRIMARY: Cereals, Citrus Fruit, Fibre Crops, Fruit, Oil Crops, Oil Crops and Cakes in Oil Equivalent, Pulses, Roots and Tubers, Sugar Crops, Treenuts and Vegetables. Data are expressed in terms of area harvested, production quantity and yield. Cereals: Area and production data on cereals relate to crops harvested for dry grain only. Cereal crops harvested for hay or harvested green for food, feed or silage or used for grazing are therefore excluded. 
2) CROPS PROCESSED: Beer of barley; Cotton lint; Cottonseed; Margarine, short; Molasses; Oil, coconut (copra); Oil, cottonseed; Oil, groundnut; Oil, linseed; Oil, maize; Oil, olive, virgin; Oil, palm; Oil, palm kernel; Oil, rapeseed; Oil, safflower; Oil, sesame; Oil, soybean; Oil, sunflower; Palm kernels; Sugar Raw Centrifugal; Wine. 
3) LIVE ANIMALS: Animals live n.e.s.; Asses; Beehives; Buffaloes; Camelids, other; Camels; Cattle; Chickens; Ducks; Geese and guinea fowls; Goats; Horses; Mules; Pigeons, other birds; Pigs; Rabbits and hares; Rodents, other; Sheep; Turkeys. 
4) LIVESTOCK PRIMARY: Beeswax; Eggs (various types); Hides buffalo, fresh; Hides, cattle, fresh; Honey, natural; Meat (ass, bird nes, buffalo, camel, cattle, chicken, duck, game, goat, goose and guinea fowl, horse, mule, Meat nes, meat other camelids, Meat other rodents, pig, rabbit, sheep, turkey); Milk (buffalo, camel, cow, goat, sheep); Offals, nes; Silk-worm cocoons, reelable; Skins (goat, sheep); Snails, not sea; Wool, greasy. 
5) LIVESTOCK PROCESSED: Butter (of milk from sheep, goat, buffalo, cow); Cheese (of milk from goat, buffalo, sheep, cow milk); Cheese of skimmed cow milk; Cream fresh; Ghee (cow and buffalo milk); Lard; Milk (dry buttermilk, skimmed condensed, skimmed cow, skimmed dried, skimmed evaporated, whole condensed, whole dried, whole evaporated); Silk raw; Tallow; Whey (condensed and dry); Yoghurt.

Production data units from FAO match the Kastner production data.

Compare format for recently downloaded FAO production data and Kastner downloaded FAO production data. Kastner provides multiple years of data while I only downloaded the 2017 production data since bulk download was not an option. Based on the variable names, it looks like Kastner did the bulk download and then separated the data into primary crops, processed crops, and primary livestock. That means we need to do a couple data formatting things to make our variable names match Kastner's.

2017 production data from Kastner to be able to match the FAO data
```{r}
# read in Kastner production crops data
k_production_crops <- fread(paste0(kastner_data_path, "Production_Crops_E_All_Data_(Normalized).csv"))
# subset to 2017
k_production_crops_17 <- k_production_crops %>% 
  filter(Year == 2017) %>% 
  head()

rm(k_production_crops)

# read in Kastner production crops processed data
k_production_crops_processed <- fread(paste0(kastner_data_path, "Production_CropsProcessed_E_All_Data_(Normalized).csv"))
# Data only available to 2014
# subset to 2014 (smaller dataset so easier to work with)
k_production_crops_processed_14 <- k_production_crops_processed %>% 
  filter(Year == 2014) %>% 
  head()

rm(k_production_crops_processed)

# read in Kastner production livestock primary data
k_production_livestock_primary <- fread(paste0(kastner_data_path, "Production_LivestockPrimary_E_All_Data_(Normalized).csv"))
# subset to 2017
k_production_livestock_primary_17 <- k_production_livestock_primary %>% 
  filter(Year == 2017) %>% 
  head()

rm(k_production_livestock_primary)
```

Function to pull individual production year data for each category (crops, processed crops, livestock)
```{r}
fao_production_cleaning <- function(data_folder, year){
  production_data <- fread(paste0(data_folder, "Production_Crops_", year, "_E_All_Data_(Normalized)_7_20_22.csv"))
}
```

#### New Data
```{r}
fao_production_crops_raw_new <- fread(paste0(production_data_path, "Production_Crops_2015_2019_FAOSTAT_6_29_2023.csv"))

test_2017 <- fao_production_crops_raw_new %>% 
  filter(Year == 2017)

compare <- fao_production_crops_raw %>% 
  filter(Element == "Production" | Element == "Yield")
```


#### Old Method
2017 FAO data - Production Crops
```{r}
# read in recently downloaded production crops data from FAO website
fao_production_crops_raw <- fread(paste0(production_data_path, "Production_Crops_2017_E_All_Data_(Normalized)_7_20_22.csv"))

# format data
production_crops <- fao_production_crops_raw %>% 
  select(-c(`Domain Code`, Domain)) %>%                                         #remove first 2 columns to match Kastner
  rename(`Area Code` = `Area Code (FAO)`) %>%                                   #rename variables to match Kastner
  rename(`Item Code` = `Item Code (FAO)`) %>% 
  relocate(c(`Item Code`, Item), .before = `Element Code`)                      #change variable order to match Kastner

# final col in our data is flag description, in Kastner it is notes. We'll keep both in for now.
write_csv(production_crops, paste0(save_path, "Production_Crops_2017_E_All_Data_(Normalized)_7_20_22_formatted.csv"))
```

2017 FAO data - Production Processed Crops
```{r}
# read in recently downloaded production crops processed data from FAO website
fao_production_crops_processed_raw <- fread(paste0(production_data_path, "Production_CropsProcessed_2017_E_All_Data_(Normalized)_7_20_22.csv"))

# format data
production_crops_processed <- fao_production_crops_processed_raw %>% 
  select(-c(`Domain Code`, Domain)) %>%                                         #remove first 2 columns to match Kastner
  rename(`Area Code` = `Area Code (FAO)`) %>%                                   # rename variables to match Kastner
  rename(`Item Code` = `Item Code (FAO)`)  %>% 
  relocate(c(`Item Code`, Item), .before = `Element Code`) %>%                  # change variable order to match Kastner
  select(-`Flag Description`)                                                   # final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.

write_csv(production_crops_processed, paste0(save_path, "Production_CropsProcessed_2017_E_All_Data_(Normalized)_7_20_22_formatted.csv"))
```

2017 FAO data - Production Livestock
```{r}
#read in recently downloaded production livestock data from FAO website
fao_production_livestock_primary_raw <- fread(paste0(fao_data_path, "Production_LivestockPrimary_2017_E_All_Data_(Normalized)_7_20_22.csv"))

# format data
production_livestock_primary <- fao_production_livestock_primary_raw %>% 
  select(-c(`Domain Code`, Domain)) %>%                                         #remove first 2 columns to match Kastner
  rename(`Area Code` = `Area Code (FAO)`) %>%                                   #rename variables to match Kastner
  rename(`Item Code` = `Item Code (FAO)`) %>% 
  relocate(c(`Item Code`, Item), .before = `Element Code`) %>% 
  select(-`Flag Description`)                                                   #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# Item codes are slightly different for some rows, but these are going to be subset just to tonnes anyways
# and the rows with tonnes as the unit all look good.

write_csv(production_livestock_primary, paste0(save_path, "Production_LivestockPrimary_2017_E_All_Data_(Normalized)_7_20_22_formatted.csv"))
```

#### Alternative method?
It looks like Haley was able to bulk download all of the years from 2015-2019 into one table for production of crops, processed crops, and livestock. A 
Read in 2015 - 2019 production data downloaded from FAO website
```{r}
# # all crops (primary and procesed)
h_production_crops <- fread("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_production_data/Production_Crops_2015_2019_E_All_Data_(Normalized)_dl_11_28_22.csv") %>% 
  filter(Year == 2017)

# # processed crops only
 h_production_crops_processed <- fread("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_production_data/Production_CropsProcessed_2015_2019_E_All_Data_(Normalized)_dl_11_28_22.csv") %>% 
  filter(Year == 2017)
 
# livestock - primary
h_production_livestock_primary <- fread("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_production_data/Production_LivestockPrimary_2015_2019_E_All_Data_(Normalized)_dl_11_28_22.csv")
h_production_livestock_primary$`Item Code (CPC)` <- as.character(h_production_livestock_primary$`Item Code (CPC)`)

# note for crops, item code and area code parenthetical are different than 2017 download
# previously we had item code in FAO units but now we have item code in CPC units
# read in item code definitions from FAO
item_code <- read_csv("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_production_data/FAO_item_code_defintions_dl_11_28_22.csv")
item_code_1 <- item_code %>%
  select(`Item Code`, `CPC Code`)
```


2015 production data
```{r}
# h_production_crops_2 <- h_production_crops %>% 
#   filter(Year == 2015) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) #change variable order to match Kastner
# # final col in our data is flag description, in Kastner it is notes. We'll keep both in for now.
# write_csv(h_production_crops_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2015/Production_Crops_2015_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_crops_processed_2 <- h_production_crops_processed %>% 
#   filter(Year == 2015) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) # final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_crops_processed_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2015/Production_CropsProcessed_2015_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_livestock_primary_2 <- h_production_livestock_primary %>% 
#   filter(Year == 2015) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% 
#   select(-`Flag Description`) #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_livestock_primary_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2015/Production_LivestockPrimary_2015_E_All_Data_(Normalized)_11_28_22_formatted.csv")
```


2016 production data
```{r}
# h_production_crops_2 <- h_production_crops %>% 
#   filter(Year == 2016) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) #change variable order to match Kastner
# # final col in our data is flag description, in Kastner it is notes. We'll keep both in for now.
# write_csv(h_production_crops_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2016/Production_Crops_2016_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_crops_processed_2 <- h_production_crops_processed %>% 
#   filter(Year == 2016) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) # final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_crops_processed_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2016/Production_CropsProcessed_2016_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_livestock_primary_2 <- h_production_livestock_primary %>% 
#   filter(Year == 2016) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% 
#   select(-`Flag Description`) #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_livestock_primary_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2016/Production_LivestockPrimary_2016_E_All_Data_(Normalized)_11_28_22_formatted.csv")
```


2018 production data
```{r}
# h_production_crops_2 <- h_production_crops %>% 
#   filter(Year == 2018) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) #change variable order to match Kastner
# # final col in our data is flag description, in Kastner it is notes. We'll keep both in for now.
# write_csv(h_production_crops_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2018/Production_Crops_2018_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_crops_processed_2 <- h_production_crops_processed %>% 
#   filter(Year == 2018) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) # final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_crops_processed_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2018/Production_CropsProcessed_2018_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_livestock_primary_2 <- h_production_livestock_primary %>% 
#   filter(Year == 2018) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% 
#   select(-`Flag Description`) #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_livestock_primary_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2018/Production_LivestockPrimary_2018_E_All_Data_(Normalized)_11_28_22_formatted.csv")
```


2019 production data
```{r}
# h_production_crops_2 <- h_production_crops %>% 
#   filter(Year == 2019) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) #change variable order to match Kastner
# # final col in our data is flag description, in Kastner it is notes. We'll keep both in for now.
# write_csv(h_production_crops_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2019/Production_Crops_2019_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_crops_processed_2 <- h_production_crops_processed %>% 
#   filter(Year == 2019) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) # final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_crops_processed_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2019/Production_CropsProcessed_2019_E_All_Data_(Normalized)_11_28_22_formatted.csv")
# 
# h_production_livestock_primary_2 <- h_production_livestock_primary %>% 
#   filter(Year == 2019) %>%
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% 
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% 
#   select(-`Flag Description`) #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
# write_csv(h_production_livestock_primary_2, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2019/Production_LivestockPrimary_2019_E_All_Data_(Normalized)_11_28_22_formatted.csv")
```



### Food balances (Kastner calls these Commodity balances, but they are definitely the FAO food balances)

There are differences in values for food balances between the old and new methodology. However, the data seem to be displayed in the same format with the same variables, so it should be fine to use the new methodology for 2017 with the Kastner script.

Food supply quantity (tonnes) in Kastner == Food in FAO data.

Kastner has detailed data for all items displayed in tonnes rather than in 1000 tonnes. I'm not sure how they got this, because I can only get these data in 1,000 tonnes. I've tried downloading just the 2017 data from the FAO website, doing a bulk download from the FAO website, and using the old (2013) and new methodology.

Downloading food balances for crops from FAO website to match Kastner data
https://www.fao.org/faostat/en/#data/FBS
-select all countries
-select elements: domestic supply quantity, export quantity, feed, food, import quantity, losses, other uses (non-food), processed, production quantity, seed, stock variation
-select all crop and processed items (includes infant food, miscellaneous, alcohol (non-food), etc.) that are not animal products (meat, eggs, aquatic plants, aquatic products, wool, silk, honey, etc.)
-select 2017

To download commodity balances for livestock and fish, select the same conditions as above, but the opposite items (select animal products but not crops).

Compare commodity balances FAO data - Kastner vs. what I just downloaded.
Convert values to tonnes (currently in 1,000 tonnes).

Food balance data - Kastner data to match FAO data with
```{r}
#read in Kastner commodity balances crops data
k_comm_bal_crops <- fread(paste0(kastner_data_path, "CommodityBalances_Crops_E_All_Data_(Normalized).csv"))
#subset to 2013 (most recent data available)
k_comm_bal_crops_13 <- k_comm_bal_crops %>% 
  filter(Year == 2013) %>% 
  head()

rm(k_comm_bal_crops)

#read in Kastner commodity balances fish and livestock data
k_comm_bal_livestock_fish <- fread(paste0(kastner_data_path, "CommodityBalances_LivestockFish_E_All_Data_(Normalized).csv"))
k_comm_bal_livestock_fish_13 <- k_comm_bal_livestock_fish %>% 
  filter(Year == 2013) %>% 
  head()

rm(k_comm_bal_livestock_fish)
```

2017 crops commodity balance data to match with Kastner formatting
```{r}
#read in recently downloaded commodity balances crops data from FAO website
comm_bal_crops_raw <- fread(paste0(fao_data_path, "CommodityBalances_Crops_2017_E_All_Data_(Normalized)_7_20_22.csv"))

comm_bal_crops <- comm_bal_crops_raw %>% 
  select(-c(`Domain Code`, Domain)) %>%                                         #remove first 2 columns to match Kastner
  rename(`Area Code` = `Area Code (FAO)`) %>%                                   #rename variables to match Kastner
  relocate(c(`Item Code`, Item), .before = `Element Code`) %>%                  #change variable order to match Kastner
  select(-`Flag Description`)                                                   #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.

#check units 
table(comm_bal_crops$Unit) #in 1000 tonnes
table(k_comm_bal_crops_13$Unit) #in tonnes

#multiply values by 1000 in h_comm_bal_crops_1 and update units
comm_bal_crops <- comm_bal_crops %>% 
  mutate(Value = Value*1000) %>% 
  mutate(Unit = "tonnes")
write_csv(comm_bal_crops, paste0(save_path, "CommodityBalances_Crops_2017_E_All_Data_(Normalized)_7_20_22_formatted.csv"))
```

2017 animals commodity balance data to match with Kastner formatting
```{r}
#read in recently downloaded commodity balances livestock data from FAO website
comm_bal_livestock_fish_raw <- fread(paste0(fao_data_path, "CommodityBalances_LivestockFish_2017_E_All_Data_(Normalized)_7_20_22.csv"))

comm_bal_livestock_fish <- comm_bal_livestock_fish_raw %>% 
  select(-c(1:2)) %>%                                                           #remove first 2 columns to match Kastner
  rename(`Area Code` = `Area Code (FAO)`) %>%                                   #rename variables to match Kastner
  relocate(c(`Item Code`, Item), .before = `Element Code`) %>%                  #change variable order to match Kastner
  
  select(-`Flag Description`)

#check units 
table(comm_bal_livestock_fish$Unit) #in 1000 tonnes
table(k_comm_bal_livestock_fish_13$Unit) #in tonnes

#multiply values by 1000 and update units
comm_bal_livestock_fish <- comm_bal_livestock_fish %>% 
  mutate(Value = Value*1000) %>% 
  mutate(Unit = "tonnes")
write_csv(comm_bal_livestock_fish, paste0(save_path, "CommodityBalances_LivestockFish_2017_E_All_Data_(Normalized)_7_20_22_formatted.csv"))
```


Read in 2015 - 2019 food balance data
```{r}
# # crops - had to download in 2 chunks
# h_fb_crops_15_16 <- fread("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_food_balance_data/CommodityBalances_Crops_2015_2016_E_All_Data_(Normalized)_dl_11_28_22.csv")
# h_fb_crops_17_19 <- fread("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_food_balance_data/CommodityBalances_Crops_2017_2019_E_All_Data_(Normalized)_dl_11_28_22.csv")
# h_fb_crops <- rbind(h_fb_crops_15_16, h_fb_crops_17_19)
# 
# # livestock
# h_fb_livestock <- fread("/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/FAO_food_balance_data/CommodityBalances_LivestockFish_2015_2019_E_All_Data_(Normalized)_dl_11_28_22.csv")
```


2015 food balance data
```{r}
# # crops
# # format data
# h_fb_crops_1 <- h_fb_crops %>% 
#   filter(Year == 2015) %>% # subset to 2015
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>%  #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_crops_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2015/CommodityBalances_Crops_2015_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
# 
# # livestock
# # format data
# h_fb_livestock_1 <- h_fb_livestock %>% 
#   filter(Year == 2015) %>% # subset to 2015
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>% 
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_livestock_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2015/CommodityBalances_LivestockFish_2015_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
```


2016 food balance data
```{r}
# # crops
# # format data
# h_fb_crops_1 <- h_fb_crops %>% 
#   filter(Year == 2016) %>% # subset to 2016
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>%  #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_crops_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2016/CommodityBalances_Crops_2016_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
# 
# # livestock
# # format data
# h_fb_livestock_1 <- h_fb_livestock %>% 
#   filter(Year == 2016) %>% # subset to 2016
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>% 
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_livestock_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2016/CommodityBalances_LivestockFish_2016_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
```


2018 food balance data
```{r}
# # crops
# # format data
# h_fb_crops_1 <- h_fb_crops %>% 
#   filter(Year == 2018) %>% # subset to 2018
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>%  #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_crops_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2018/CommodityBalances_Crops_2018_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
# 
# # livestock
# # format data
# h_fb_livestock_1 <- h_fb_livestock %>% 
#   filter(Year == 2018) %>% # subset to 2018
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>% 
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_livestock_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2018/CommodityBalances_LivestockFish_2018_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
```


2019 food balance data
```{r}
# # crops
# # format data
# h_fb_crops_1 <- h_fb_crops %>% 
#   filter(Year == 2019) %>% # subset to 2019
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>%  #final col in our data is flag description, Kastner does not have a final column 'notes' so we'll delete the flag description column.
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_crops_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2019/CommodityBalances_Crops_2019_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
# 
# # livestock
# # format data
# h_fb_livestock_1 <- h_fb_livestock %>% 
#   filter(Year == 2019) %>% # subset to 2019
#   left_join(item_code_1,  by = c("Item Code (CPC)" = "CPC Code")) %>% # join with item codes df to get FAO item codes
#   select(-c(1:2, 7)) %>% #remove first 2 columns to match Kastner
#   rename(`Area Code` = `Area Code (M49)`) %>% #rename variables to match Kastner
#   relocate(c(`Item Code`, Item), .before = `Element Code`) %>% #change variable order to match Kastner
#   select(-`Flag Description`) %>% 
#   mutate(Value = Value*1000) %>% #multiply values by 1000 and update units
#   mutate(Unit = "tonnes")
# # save csv
# write_csv(h_fb_livestock_1, "/home/shares/food-systems/social_justice_projects/food_trade_ej/kastner_2022_data_code/Data_2019/CommodityBalances_LivestockFish_2019_E_All_Data_(Normalized)_dl_11_28_22_formatted.csv")
```


