---
title: "01a-format-fao-trade"
date: "2023-05-09"
editor_options: 
  chunk_output_type: console
---

Objective: Format the 2017 FAO production, commodity balances, and detailed trade matrix datasets to match Kastner's data that is input into the trade matrix script.
 
Update: Formatted data for 2015, 2016, 2017, 2018, and 2019 on June 29, 2023.

```{r}
library(tidyverse)
library(data.table)
```


### Setup
```{r}
raw_folder_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
kastner_data_path <- paste0(raw_folder_path, "kastner_2022/")
fao_data_path <- paste0(raw_folder_path, "FAO/2017_FAO_data/")
production_data_path <- paste0(raw_folder_path, "FAO/production_data/")
fb_data_path <- paste0(raw_folder_path, "FAO/food_balance_sheets/")
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01a_format_fao_trade/"
```

### Detailed trade matrix

Bulk download of trade matrix data was used and subset to years of interest.

We need to gapfill missing trade data with trading partners (FAO does this for some, but not all).

Kastner Trade Matrix example to understand how to format the trade data.
```{r}
#read in Kastner detailed trade matrix data (from FAO) - This is to understand what we want the output data to look like
k_trade_matrix <- fread(paste0(kastner_data_path, "Trade_DetailedTradeMatrix_E_All_Data_(Normalized).csv"))
head_k_trade_matrix <- head(k_trade_matrix)
# k_trade_matrix is not used anywhere else
rm(k_trade_matrix)
```

Trade data - Gapfill imports with export data
```{r}
#read in recently downloaded FAO detailed trade matrix data
fao_trade_matrix_raw <- fread(paste0(fao_data_path, "Trade_DetailedTradeMatrix_E_All_Data_(Normalized)_7_14_22.csv"))

# Function to subset trade matrix by year of interest
fao_trade_quantity_gapfill <- function(fao_trade_matrix_raw, year) {
  # subset columns to match Kastner data columns
  trade_matrix <- fao_trade_matrix_raw %>% 
  select(-c(`Reporter Country Code (M49)`, `Partner Country Code (M49)`, `Item Code (CPC)`)) %>% 
  filter(Year == year)
  
  # want to prioritize import data and only gapfill with export data
  # from Kastner paper:
  # "There are, for example, sometimes differences between a reported import from country A importing from country B, and the corresponding report of an export from country B to country A. There are also sometimes trade flows missing from reports (Gehlhar 1996). To harmonize these reports, we decided to prioritize import reports as being more reliable and only used reported exports to fill missing flows. This is a reasonable assumption and feasible for such a huge dataset."
  
  # separate import and export data into separate dataframes
  # rename export data so it can be merged with import data, make the columns match up and switch the reporter and partner country names
  export <- trade_matrix %>% 
    filter(`Element Code` == 5910) %>% 
    rename("Partner Countries_1" = "Reporter Countries") %>%
    rename("Partner Country Code_1" = "Reporter Country Code") %>% 
    rename("Reporter Countries" = "Partner Countries") %>% 
    rename("Reporter Country Code" = "Partner Country Code") %>% 
    rename("Partner Countries" = "Partner Countries_1") %>% 
    rename("Partner Country Code" = "Partner Country Code_1") %>% 
    rename("Value_gapfill" = Value) %>% 
    rename("Flag_gapfill" = Flag) %>% 
    select(-c("Element Code", "Element"))
  
  import <- trade_matrix %>% 
    filter(`Element Code` == 5610)

  # join import and export data back together 
  import_export <- full_join(import, export, by = c("Reporter Countries", "Reporter Country Code", "Partner Countries", "Partner Country Code", "Item Code", "Item", "Year Code", "Year", "Unit"))
  
  # fill in missing import quantities with export quantities and make note in Flag column where that's done ("gapfill")
  trade_df_gapfilled <- import_export %>% 
    mutate(Value = ifelse(is.na(Value), Value_gapfill, Value)) %>% 
    mutate(Flag = ifelse(is.na(Flag), "gapfill", Flag)) %>% 
    mutate(`Element Code` = ifelse(is.na(`Element Code`), 5610, `Element Code`)) %>% 
    mutate(Element = ifelse(is.na(Element), "Import Quantity", Element)) %>% 
    select(-c(Value_gapfill, Flag_gapfill))
  
  return(trade_df_gapfilled)
}
```


Gapfill and save as csv for 2015, 2016, 2017, 2018, 2019 data
```{r}
# 2015 data
trade_2015 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2015)

# save df - can take out NAs later
write_csv(trade_2015, paste0(save_path,"Trade_2015_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2016 data
trade_2016 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2016)

# save df - can take out NAs later
write_csv(trade_2016, paste0(save_path,"Trade_2016_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2017 data
trade_2017 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2017)

# save df - can take out NAs later
write_csv(trade_2017, paste0(save_path,"Trade_2017_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2018 data
trade_2018 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2018)

# save df - can take out NAs later
write_csv(trade_2018, paste0(save_path,"Trade_2018_DetailedTradeMatrix_Gapfilled_formatted.csv"))

# 2019 data
trade_2019 <- fao_trade_quantity_gapfill(fao_trade_matrix_raw = fao_trade_matrix_raw,
                                         year = 2019)

# save df - can take out NAs later
write_csv(trade_2019, paste0(save_path,"Trade_2019_DetailedTradeMatrix_Gapfilled_formatted.csv"))
```

Average trade across five years
```{r}
# get the unique values of this row to add back in later
unique(trade_2016$`Element Code`)
unique(trade_2016$Element)
unique(trade_2016$Unit)

# rbind trade df for five years of interest
average_trade <- rbind(trade_2015,
                       trade_2016,
                       trade_2017,
                       trade_2018,
                       trade_2019) %>%
  # group by all columns of specificity except for year and description columns
  group_by(`Reporter Country Code`,
           `Reporter Countries`,
           `Partner Country Code`,
           `Partner Countries`,
           `Item Code`,
           Item) %>% 
  summarise(Value_average = mean(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # add back in three columns that are missing when groupby was done
  # after adding these columns back everything should match the columns of the Kastner matrix
  mutate(`Element Code` = 5610,
         Element = "Import Quantity",
         `Year Code` = "2015-2019",
         Year = "2015-2019",
         Unit = "tonnes",
         Flag = NA) %>% 
  # relocate columns to match order of kastner matrix
  relocate(Value_average, .after = Unit)

# Test where the NaN are coming from in the Value_average column
NaN_test <- average_trade %>%  
  filter(is.nan(Value_average))
# NaN only occur in rows where the Value is NA so there is nothing we can really do about this
# Also only 20,000 rows out of almost two million observationsare NaN so will ignore and filter for these later

write_csv(average_trade, paste0(save_path,"Trade_average_DetailedTradeMatrix_Gapfilled_formatted.csv"))

rm(fao_trade_matrix_raw,
   trade_2015,
   trade_2016,
   trade_2017,
   trade_2018,
   trade_2019,
   NaN_test)
```


### Production data

Data in the bulk download files is shown pivoted by years (wide format), except for the Normalized file which is in database format (long format).

Kastner FAO production data is normalized data (long format) separated into crops, processed crops, and primary livestock. Downloading selected options from the FAO website provides the data in long format (i.e. normalized).

Downloading production data from FAO website to match Kastner data: https://www.fao.org/faostat/en/#data/QCL
- From "Bulk Downloads" select "All Data Normalized"
- .csv file labeled `Production_Crops_Livestock_E_All_Data_(Normalized)` is our file of interest

FAO products included in primary crops, processed crops, primary livestock, and processed livestock.
Crop and livestock statistics are recorded for 278 products, covering the following categories: 
1) CROPS PRIMARY: Cereals, Citrus Fruit, Fibre Crops, Fruit, Oil Crops, Oil Crops and Cakes in Oil Equivalent, Pulses, Roots and Tubers, Sugar Crops, Treenuts and Vegetables. Data are expressed in terms of area harvested, production quantity and yield. Cereals: Area and production data on cereals relate to crops harvested for dry grain only. Cereal crops harvested for hay or harvested green for food, feed or silage or used for grazing are therefore excluded. 
2) CROPS PROCESSED: Beer of barley; Cotton lint; Cottonseed; Margarine, short; Molasses; Oil, coconut (copra); Oil, cottonseed; Oil, groundnut; Oil, linseed; Oil, maize; Oil, olive, virgin; Oil, palm; Oil, palm kernel; Oil, rapeseed; Oil, safflower; Oil, sesame; Oil, soybean; Oil, sunflower; Palm kernels; Sugar Raw Centrifugal; Wine. 
3) LIVE ANIMALS: Animals live n.e.s.; Asses; Beehives; Buffaloes; Camelids, other; Camels; Cattle; Chickens; Ducks; Geese and guinea fowls; Goats; Horses; Mules; Pigeons, other birds; Pigs; Rabbits and hares; Rodents, other; Sheep; Turkeys. 
4) LIVESTOCK PRIMARY: Beeswax; Eggs (various types); Hides buffalo, fresh; Hides, cattle, fresh; Honey, natural; Meat (ass, bird nes, buffalo, camel, cattle, chicken, duck, game, goat, goose and guinea fowl, horse, mule, Meat nes, meat other camelids, Meat other rodents, pig, rabbit, sheep, turkey); Milk (buffalo, camel, cow, goat, sheep); Offals, nes; Silk-worm cocoons, reelable; Skins (goat, sheep); Snails, not sea; Wool, greasy. 
5) LIVESTOCK PROCESSED: Butter (of milk from sheep, goat, buffalo, cow); Cheese (of milk from goat, buffalo, sheep, cow milk); Cheese of skimmed cow milk; Cream fresh; Ghee (cow and buffalo milk); Lard; Milk (dry buttermilk, skimmed condensed, skimmed cow, skimmed dried, skimmed evaporated, whole condensed, whole dried, whole evaporated); Silk raw; Tallow; Whey (condensed and dry); Yoghurt.

Production data units from FAO match the Kastner production data.

Based on the variable names, it looks like Kastner did the bulk download and then separated the data into primary crops, processed crops, and primary livestock. We will replicate this process below and subset for our years of interest.

2017 production data from Kastner to be able to match the FAO data
```{r}
# read in Kastner production crops data
k_production_crops <- fread(paste0(kastner_data_path, "Production_Crops_E_All_Data_(Normalized).csv"))
# subset to 2017
k_production_crops_17 <- k_production_crops %>% 
  filter(Year == 2017) %>% 
  head()

rm(k_production_crops)

# read in Kastner production crops processed data
k_production_crops_processed <- fread(paste0(kastner_data_path, "Production_CropsProcessed_E_All_Data_(Normalized).csv"))
# Data only available to 2014
# subset to 2014 (smaller dataset so easier to work with)
k_production_crops_processed_14 <- k_production_crops_processed %>% 
  filter(Year == 2014) %>% 
  head()

rm(k_production_crops_processed)

# read in Kastner production livestock primary data
k_production_livestock_primary <- fread(paste0(kastner_data_path, "Production_LivestockPrimary_E_All_Data_(Normalized).csv"))
# subset to 2017
k_production_livestock_primary_17 <- k_production_livestock_primary %>% 
  filter(Year == 2017) %>% 
  head()

rm(k_production_livestock_primary)
```

#### New Data
All Data Bulk Download
```{r}
all_production_raw <- fread(paste0(production_data_path, "Production_Crops_Livestock_E_All_Data_(Normalized)_06_29_2023.csv"))

# clean raw data to simplify
all_production <- all_production_raw %>% 
  # Filter only for years of interest
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019)) %>% 
  # remove extra area code and item code columns, remaining are in FAO format
  select(-`Area Code (M49)`, - `Item Code (CPC)`) %>% 
  # Only need production data in tonnes so filter for element code 5510
  filter(`Element Code` == 5510,
         # remove regions and China aggregated areas
         `Area Code` <= 300)

# Take average over years 
all_production_average <- all_production %>% 
  # group by all columns of specificity except for year and description columns
  group_by(`Area Code`,
           Area,
           `Item Code`,
           Item,
           `Element Code`,
           Element) %>% 
  # Value_average is the mean of the value column for that group
  summarise(Value_average = mean(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # add back in three columns that are missing when groupby was done
  # after adding these columns back everything should match the columns of the Kastner crop production data
  mutate(`Year Code` = "2015-2019",
         Year = "2015-2019",
         Unit = "tonnes",
         Flag = NA) %>% 
  relocate(Value_average, .after = Unit)

units_check <- all_production %>% 
  distinct(`Element Code`, Element, Unit)

production_items <- all_production %>% 
  distinct(`Item Code`, Item)
```

Make lists of products for crops, crops processed, and livestock primary based on the "Items" groupings from FAO. Downloaded these from the FAO website.
```{r}
crops_list <- fread(paste0(production_data_path, "Production_Crops_2015_2019_FAOSTAT_6_29_2023.csv")) %>% 
  distinct(`Item Code (FAO)`, Item)

crops_processed_list <- fread(paste0(production_data_path, "Production_CropsProcessed_2015_2019_FAOSTAT_6_29_2023.csv")) %>% 
  distinct(`Item Code (FAO)`, Item)

livestock_primary_list <- fread(paste0(production_data_path, "Production_LivestockPrimary_2015_2019_FAOSTAT_6_29_2023.csv")) %>% 
  distinct(`Item Code (FAO)`, Item)
```

Separate `all_production_average` by the appropriate groupings of crops, crops processed, and livestock primary.
```{r}
# separate crops production data
crops_production_average <- all_production_average %>% 
  filter(`Item Code` %in% c(crops_list$`Item Code (FAO)`))

# save crops production data 
write_csv(crops_production_average, paste0(save_path, "Production_Crops_2015_2019_FAOSTAT_6_29_2023_formatted.csv"))

# separate processed crops production data
processed_crops_production_average <- all_production_average %>% 
  filter(`Item Code` %in% c(crops_processed_list$`Item Code (FAO)`))

# save crops production data 
write_csv(processed_crops_production_average, paste0(save_path, "Production_CropsProcessed_2015_2019_FAOSTAT_6_29_2023_formatted.csv"))

# separate processed crops production data
livestock_primary_production_average <- all_production_average %>% 
  filter(`Item Code` %in% c(livestock_primary_list$`Item Code (FAO)`))

# save crops production data 
write_csv(livestock_primary_production_average, paste0(save_path, "Production_LivestockPrimary_2015_2019_FAOSTAT_6_29_2023_formatted.csv"))
```

### Food balances (Kastner calls these Commodity balances, but they are definitely the FAO food balances)

There are differences in values for food balances between the old and new methodology. However, the data seem to be displayed in the same format with the same variables, so it should be fine to use the new methodology for our years of interest with the Kastner script.

Food supply quantity (tonnes) in Kastner == Food in FAO data.

Kastner has detailed data for all items displayed in tonnes rather than in 1000 tonnes. When downloading data from FAO the units are 1000 tonnes. To fix this we will just multiply by 1000 and change the unit label. We will use the bulk download data from the FAO website and clean it to match the Kastner data and suit it to our needs.

Downloading food balances for crops and livestock from FAO website to match Kastner data: https://www.fao.org/faostat/en/#data/FBS
- From "Bulk Downloads" select "All Data Normalized"
- .csv file labeled `FoodBalanceSheets_E_All_Data_(Normalized)` is our file of interest

The bulk download will need to be cleaned to match the Kastner data. Convert values to tonnes (currently in 1,000 tonnes). Based on the Kastner script the Food Balance Data is used to fill in offals and sugars data so we only need element codes 5511 for Production (offals) and 5131 for Processing (sugar). Based on this only these element codes are being kept and then averaged. If this needs to be changed it will be done and the notes updated here.

Food balance data - Kastner data to match FAO data with
```{r}
#read in Kastner commodity balances crops data
k_comm_bal_crops <- fread(paste0(kastner_data_path, "CommodityBalances_Crops_E_All_Data_(Normalized).csv"))
#subset to 2013 (most recent data available)
k_comm_bal_crops_13 <- k_comm_bal_crops %>% 
  filter(Year == 2013) %>% 
  head()

# # Make list of crop items from FAO included in Kastner commodity balances
# k_comm_bal_crops_items <- k_comm_bal_crops %>% 
#   distinct(`Item Code`, Item)
# 
# # Make list of crop elements from FAO included in Kastner commodity balances
# k_comm_bal_crops_elements <- k_comm_bal_crops %>% 
#   distinct(`Element Code`, Element)

rm(k_comm_bal_crops)

#read in Kastner commodity balances fish and livestock data
k_comm_bal_livestock_fish <- fread(paste0(kastner_data_path, "CommodityBalances_LivestockFish_E_All_Data_(Normalized).csv"))
k_comm_bal_livestock_fish_13 <- k_comm_bal_livestock_fish %>% 
  filter(Year == 2013) %>% 
  head()

# # Make list of livestock and fish items from FAO included in Kastner commodity balances
# k_comm_bal_livestock_fish_items <- k_comm_bal_livestock_fish %>% 
#   distinct(`Item Code`, Item)
# 
# # Make list of livestock and fish elements from FAO included in Kastner commodity balances
# # This list should be the exact same as the crop list
# k_comm_bal_livestock_fish_elements <- k_comm_bal_livestock_fish %>% 
#   distinct(`Element Code`, Element) %>% 
#   arrange(`Element Code`)

rm(k_comm_bal_livestock_fish)
```

Clean Food Balance bulk download and clean to match Kastner data format. Take averages across years.
```{r}
all_fb_raw <- fread(paste0(fb_data_path, "FoodBalanceSheets_E_All_Data_(Normalized)_06_29_2023.csv"))

element_list <- all_fb_raw %>% 
  distinct(`Element Code`, Element)

# Clean and remove excess data from bulk download
all_fb <- all_fb_raw %>% 
  # Filter only for years of interest
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019)) %>% 
  # remove extra area code and item code columns, remaining are in FAO format
  select(-`Area Code (M49)`, - `Item Code (CPC)`) %>% 
  # remove rows with item code 2501 because it is population
  filter(`Item Code` != 2501,
         `Element Code` %in% c(5511, 5131)) %>% 
  # multiply values by 1000 in and update units label to match Kastner
  mutate(Value = Value*1000) %>% 
  mutate(Unit = "tonnes")

# Take average over years 
commodity_balance_averages <- all_fb %>% 
  # group by all columns of specificity except for year and description columns
  group_by(`Area Code`,
           Area,
           `Item Code`,
           Item,
           `Element Code`,
           Element) %>% 
  # Value_average is the mean of the value column for that group
  summarise(Value_average = mean(Value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # add back in three columns that are missing when groupby was done
  # after adding these columns back everything should match the columns of the Kastner crop production data
  mutate(`Year Code` = "2015-2019",
         Year = "2015-2019",
         Unit = "tonnes",
         Flag = NA) %>% 
  relocate(Value_average, .after = Unit)

write_csv(commodity_balance_averages, paste0(save_path, "CommodityBalances_2015_2019_(Normalized)_06_29_2023_formatted.csv"))

```

