---
title: "01c_allocate_unknown_trade_fish_consumption"
output: html_document
date: "2023-05-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Similar to the other fish trade, fish oil and fish meal (FOFM) trade had unknown source countries. Following the same method as discussed above, we proportionately bumped up all known FOFM traded to a country from the unknown source trade. 

## Setup
```{r}
library(tidyverse)
```

### File paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
matrix_path <- paste0(raw_data_path, "trade/aquatic_food_trade/")
production_path <- paste0(raw_data_path, "FAO/production_data/FI_Trade_PP_2023.1.0/")
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01d_allocate_unknown_trade_fmfo_consumption/"
```

## Read in data
```{r}
fmfo_matrix <- read.csv(paste0(matrix_path, "20230626_nceas_consumption_fmfo.csv"))

```

## Average the data over the years (2015-2019)
For each consumer country, source country, habitat, method, nceas_group (species) we want to take the average consumption_live_t over the five years of interest.
```{r}
colnames(fmfo_matrix)

# Group by everything except for year and take the mean of consumption_live_t
fmfo_matrix_mean <- fmfo_matrix %>% 
  # rename to match other trade data
  rename(consumer_iso3c = consuming_iso3c) %>% 
  group_by(consumer_iso3c, source_country_iso3c, dom_source) %>% 
  summarise(consumption_live_weight_t = mean(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup()
```

## Identify combinations of unknown data
```{r}
# Identify typologies of unknown 
unique(fmfo_matrix_mean$consumer_iso3c)
unique(fmfo_matrix_mean$source_country_iso3c)
unique(fmfo_matrix_mean$dom_source)
# it looks like there are only source country unknowns but lets double check

unknown_types <- fmfo_matrix_mean %>%
  filter(source_country_iso3c == "unknown"|
           consumer_iso3c == "unknown"|
           dom_source == "unknown") %>%
  select(source_country_iso3c, consumer_iso3c, dom_source) %>%
  mutate(
    source_country_iso3c = case_when(source_country_iso3c != "unknown" ~ "known",
                                     TRUE ~ "unknown"),
    consumer_iso3c = case_when(consumer_iso3c != "unknown" ~ "known",
                            TRUE ~ "unknown"),
    dom_source = case_when(dom_source != "unknown" ~ "known",
                        TRUE ~ "unknown")
    ) %>%
  group_by(source_country_iso3c, consumer_iso3c, dom_source) %>%
  tally()
# there are only source country unknowns

# Add typology group to source-consumption data frame
fmfo_matrix_mean <- fmfo_matrix_mean %>%
  mutate(unknown_type = case_when(source_country_iso3c == "unknown" ~ "source_unknown",
                                  TRUE ~ "known"))


# Check share of consumption originally unknown
fmfo_unknown_shares <- fmfo_matrix_mean %>%
  mutate(unknown_type = case_when(
    unknown_type == "known" ~ "known",
    unknown_type != "known" ~ "unknown"
  )) %>%
  group_by(consumer_iso3c, unknown_type) %>%
  summarise(consumption_live_weight_t = sum(consumption_live_weight_t)) %>%
  group_by(consumer_iso3c) %>%
  mutate(total = sum(consumption_live_weight_t),
         percent = 100*consumption_live_weight_t/total)

total_unknown <- fmfo_unknown_shares %>% 
  filter(unknown_type == "unknown") %>% 
  group_by(unknown_type) %>% 
  summarise(total_unknown = sum(consumption_live_weight_t),
            total_all = sum(total)) %>% 
  ungroup() %>% 
  mutate(percent_unknown = (total_unknown/total_all)*100)

#4.06 % of tonnes traded are unknown
```


## We only have one type of unknown for the fofm data, where we do not know the source country 
```{r}
# Use proportions of all known consumption
source_known_prop_df <- fmfo_matrix_mean %>%
  # Remove rows where source is unknown
  filter(unknown_type != "source_unknown") %>%
  # Get proportions for each consuming country
  group_by(consumer_iso3c) %>%
  mutate(source_known_prop = consumption_live_weight_t/sum(consumption_live_weight_t)) %>% 
  ungroup()

unknowns <- fmfo_matrix_mean %>%
              # Filter to rows where source is unknown
              filter(unknown_type == "source_unknown") %>%
              select(consumer_iso3c, source_unknown_t = consumption_live_weight_t)

source_unknown <- source_known_prop_df %>% 
  # join unknown data
  left_join(fmfo_matrix_mean %>%
              # Filter to rows where source is unknown
              filter(unknown_type == "source_unknown") %>%
              select(consumer_iso3c, source_unknown_t = consumption_live_weight_t),
            by = "consumer_iso3c") %>%
  # make rows with no unknown sources 0 instead of NA
  mutate(source_unknown_t = if_else(condition = is.na(source_unknown_t),
                                    true = 0,
                                    false = source_unknown_t)) %>% 
  mutate(adjust_live_weight_t = source_known_prop*source_unknown_t)
  #select(-source_known_prop, -source_unknown_t, -unknown_type)

check <- source_unknown %>% 
  group_by(consumer_iso3c, source_unknown_t) %>% 
  summarise(adjust_live_weight_t = sum(adjust_live_weight_t)) %>% 
  mutate(diff = source_unknown_t - adjust_live_weight_t)

# The biggest difference between the adjusted tonnes and unknowns is e-11 so it is negligible and just due to R rounding


```

### Combine proportioned unknown tonnes with the consumption_live_weight_t
```{r}
# Join all together
fmfo_matrix_new <- source_unknown %>% 
  # add consumption_live_weight_t and adjust_live_weight_t
  mutate(consumption_live_weight_t = consumption_live_weight_t+adjust_live_weight_t) %>% 
  select(consumer_iso3c, source_country_iso3c, dom_source, consumption_live_weight_t)

```

### Check consumption
Check to make sure the total consumption is not different for countries.
```{r}
# take average consumption over years data frame
fmfo_check <- fmfo_matrix_mean %>%
  group_by(consumer_iso3c) %>%
  # total consumption weight
  summarise(consumption_original = sum(consumption_live_weight_t, na.rm = TRUE)) %>%
  # join with adjusted data
  left_join(fmfo_matrix_new %>%
              group_by(consumer_iso3c) %>%
              # total consumption weight
              summarise(consumption_new = sum(consumption_live_weight_t, na.rm = TRUE)),
            by = "consumer_iso3c") %>%
  mutate(diff = consumption_new-consumption_original) %>%
  mutate(perc_diff = 100*(diff/consumption_original)) %>%
  arrange(desc(abs(perc_diff)))

```

Any differences in tonnage between the the original data and new data is extremely small (e-14), this is likely due to rounding of precise numbers in R so nothing to worry about.

###  Save
```{r}

write_csv(fmfo_matrix_new, paste0(save_path, "fmfo_trade_matrix_w_unknowns_allocated.csv"))

```

## FOFM Production Data 
This section will prepare the FOFM production data as the trade data does not include internal production of FOFM products. [Need Jessica to write a better explanation as to why we need to do this step.]

The FAO commodity code for the fish oil and fishmeal products we are interested in start with "081.42". We will filter the commodity list to only include products with this code at the beginning of the code.

### Read in Production data and codes
```{r}
production_raw <- read.csv(file = paste0(production_path, "TRADE_PP_QUANTITY.csv"))

commodity_list <- read.csv(file = paste0(production_path, "CL_FI_COMMODITY_ISSCFC.csv")) %>% 
  # keep code column and English name
  select(Code, Name_En) %>% 
  # filter for our product code
  # 081.42 is for all FMFO products
  filter(str_detect(string = Code, pattern = "081.42")) %>% 
  # rename Name_en to product
  rename(product = Name_En,
         product_code = Code)

country_list <- read.csv(file = paste0(production_path, "CL_FI_COUNTRY_GROUPS.csv")) %>%
  # keep UN code, iso3c, and english name
  select(UN_Code, ISO3_Code, country_name = Name_En)
```

### Clean up production data
```{r}
production_clean <- production_raw %>% 
  # filter for 2015-2019
  filter(PERIOD %in% c(2015, 2016, 2017, 2018, 2019)) %>% 
  # rename to match commodity list
  rename(product_code = COMMODITY.FAO_CODE,
         UN_Code = COUNTRY.UN_CODE) %>% 
  left_join(y = commodity_list,
            by = "product_code") %>% 
  # filter out NAs (not our products)
  filter(!is.na(product)) %>% 
  #join country names
  left_join(y = country_list,
            by = "UN_Code") %>% 
  # remove production code
  select(-PPRODUCTION_ELEMENT.ALPHA_CODE, -STATUS) %>% 
  relocate(ISO3_Code, country_name, .before = UN_Code)
```

### Average Production data
Average production of all FOFM products over years
```{r}
production_average <- production_clean %>% 
  # first sum total production of all production across each year
  group_by(ISO3_Code, country_name, PERIOD, MEASURE) %>% 
  summarise(tonnes_annual_production = sum(VALUE)) %>% 
  ungroup() %>% 
  # next average the annual tonnes produced
  group_by(ISO3_Code, country_name, MEASURE) %>% 
  summarise(tonnes_annual_production = mean(tonnes_annual_production, na.rm = TRUE)) %>% 
  ungroup()

```

Now to convert tonnes product weight to liveweight equivalent we need to multiply the Tonnage by 2.975207 per IFFO, 2009 https://essopenarchive.org/doi/full/10.22541/essoar.167590829.99780929

```{r}
conversion_value <- 2.975207

production_average_converted <- production_average %>% 
  mutate(production_live_weight_t = tonnes_annual_production * conversion_value) %>% 
  select(-MEASURE, -tonnes_annual_production) %>% 
  filter(production_live_weight_t != 0)
  
```


## Join Production and Trade
Here we join the production and trade data to adjust it for true domestic and foreign consumption. 
```{r}
fmfo_imports <- fmfo_matrix_new %>% 
  group_by(consumer_iso3c) %>% 
  summarise(import_live_weight_t = sum(consumption_live_weight_t)) %>% 
  ungroup() 

fmfo_exports <- fmfo_matrix_new %>% 
  group_by(source_country_iso3c) %>% 
  summarise(export_live_weight_t = sum(consumption_live_weight_t)) %>% 
  ungroup() 

# full join with imports and export data
fmfo_supply <- full_join(x = production_average_converted, y = fmfo_imports,
                                by = c("ISO3_Code" = "consumer_iso3c")) %>% 
  full_join(y = fmfo_exports,
            by = c("ISO3_Code" = "source_country_iso3c")) %>% 
  select(-country_name) %>% 
  mutate(production_live_weight_t = if_else(condition = is.na(production_live_weight_t),
                                             true = 0,
                                             false = production_live_weight_t),
         import_live_weight_t = if_else(condition = is.na(import_live_weight_t),
                                             true = 0,
                                             false = import_live_weight_t),
         export_live_weight_t = if_else(condition = is.na(export_live_weight_t),
                                             true = 0,
                                             false = export_live_weight_t)) %>% 
  # calculate apparent consumption
  mutate(supply = production_live_weight_t + import_live_weight_t - export_live_weight_t) %>% 
  # only allocate for countries with positive consumption
  filter(supply > 0) %>% 
  # calculate the import ratio of total supply
  mutate(import_ratio = import_live_weight_t/(production_live_weight_t + import_live_weight_t)) %>% 
  mutate(imported_fmfo_live_weight_t = import_ratio * supply,
         domestic_fmfo_live_weight_t = (1-import_ratio) * supply)

domestic_fmfo_consumption <- fmfo_supply %>% 
  select(consumer_iso3c = ISO3_Code, consumption_live_weight_t = domestic_fmfo_live_weight_t) %>% 
  mutate(source_country_iso3c = consumer_iso3c,
         dom_source = "domestic")

source_percentages <- fmfo_matrix_new %>% 
  filter(dom_source == "foreign") %>% 
  group_by(consumer_iso3c) %>% 
  mutate(proportion_by_source = consumption_live_weight_t/sum(consumption_live_weight_t)) %>% 
  select(-consumption_live_weight_t)

foreign_fmfo_consumption <- fmfo_supply %>% 
  select(consumer_iso3c = ISO3_Code, imported_fmfo_live_weight_t) %>% 
  left_join(y = source_percentages,
            by = "consumer_iso3c") %>% 
  mutate(consumption_live_weight_t = imported_fmfo_live_weight_t * proportion_by_source) %>% 
  select(-imported_fmfo_live_weight_t, -proportion_by_source)

fmfo_consumption <- bind_rows(domestic_fmfo_consumption,
                              foreign_fmfo_consumption) %>% 
  relocate(consumption_live_weight_t, .after = dom_source)
  

```


### Save
```{r}
write_csv(fmfo_consumption, paste0(save_path, "fmfo_trade_matrix_w_unknowns_allocated.csv"))
```




