---
title: "01c_allocate_unknown_trade_fish_consumption"
output: html_document
date: "2023-05-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objective: The fish trade matrix for human consumption has a number of data points that are "unknown" in the source country, habitat, method, and nceas_group columns. Unknowns arise from the error term or cases where products move through more than 2 intermediate countries. We will proportionately bump up all marine flows with the marine-unknown volume, all inland flows with the inland-unknown, and then all aquatic food flows with the unknown-unknown. This will keep the total consumption constant without having to assume a global average for the pressure values when integrated.

Additionally, the fish meal fish oil trade data needs to averages across years and saved so that pressures can be applied.

## Setup
```{r}
library(tidyverse)
```

### File paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
matrix_path <- paste0(raw_data_path, "trade/aquatic_food_trade/")
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01e_allocate_unknown_trade_fmfo_consumption/"
```

## Read in data
```{r}
fmfo_matrix <- read.csv(paste0(matrix_path, "20230626_nceas_consumption_fmfo.csv"))

```

## Average the data over the years (2015-2019)
For each consumer country, source country, habitat, method, nceas_group (species) we want to take the average consumption_live_t over the five years of interest.
```{r}
colnames(fmfo_matrix)

# Group by everything except for year and take the mean of consumption_live_t
fmfo_matrix_mean <- fmfo_matrix %>% 
  # rename to match other trade data
  rename(consumer_iso3c = consuming_iso3c) %>% 
  group_by(consumer_iso3c, source_country_iso3c, dom_source) %>% 
  summarise(consumption_live_weight_t = mean(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup()
```

## Identify combinations of unknown data
```{r}
# Identify typologies of unknown 
unique(fmfo_matrix_mean$consumer_iso3c)
unique(fmfo_matrix_mean$source_country_iso3c)
unique(fmfo_matrix_mean$dom_source)
# it looks like there are only source country unknowns but lets double check

unknown_types <- fmfo_matrix_mean %>%
  filter(source_country_iso3c == "unknown"|
           consumer_iso3c == "unknown"|
           dom_source == "unknown") %>%
  select(source_country_iso3c, consumer_iso3c, dom_source) %>%
  mutate(
    source_country_iso3c = case_when(source_country_iso3c != "unknown" ~ "known",
                                     TRUE ~ "unknown"),
    consumer_iso3c = case_when(consumer_iso3c != "unknown" ~ "known",
                            TRUE ~ "unknown"),
    dom_source = case_when(dom_source != "unknown" ~ "known",
                        TRUE ~ "unknown")
    ) %>%
  group_by(source_country_iso3c, consumer_iso3c, dom_source) %>%
  tally()
# there are only source country unknowns

# Add typology group to source-consumption data frame
fmfo_matrix_mean <- fmfo_matrix_mean %>%
  mutate(unknown_type = case_when(source_country_iso3c == "unknown" ~ "source_unknown",
                                  TRUE ~ "known"))


# Check share of consumption originally unknown
fmfo_unknown_shares <- fmfo_matrix_mean %>%
  mutate(unknown_type = case_when(
    unknown_type == "known" ~ "known",
    unknown_type != "known" ~ "unknown"
  )) %>%
  group_by(consumer_iso3c, unknown_type) %>%
  summarise(consumption_live_weight_t = sum(consumption_live_weight_t)) %>%
  group_by(consumer_iso3c) %>%
  mutate(total = sum(consumption_live_weight_t),
         percent = 100*consumption_live_weight_t/total)
```


## We only have one type of unknown for the fofm data, where we do not know the source country 
```{r}
# Use proportions of all known consumption
source_known_prop_df <- fmfo_matrix_mean %>%
  # Remove rows where source is unknown
  filter(unknown_type != "source_unknown") %>%
  # Get proportions for each consuming country
  group_by(consumer_iso3c) %>%
  mutate(source_known_prop = consumption_live_weight_t/sum(consumption_live_weight_t)) %>% 
  ungroup()

unknowns <- fmfo_matrix_mean %>%
              # Filter to rows where source is unknown
              filter(unknown_type == "source_unknown") %>%
              select(consumer_iso3c, source_unknown_t = consumption_live_weight_t)

source_unknown <- source_known_prop_df %>% 
  # join unknown data
  left_join(fmfo_matrix_mean %>%
              # Filter to rows where source is unknown
              filter(unknown_type == "source_unknown") %>%
              select(consumer_iso3c, source_unknown_t = consumption_live_weight_t),
            by = "consumer_iso3c") %>%
  # make rows with no unknown sources 0 instead of NA
  mutate(source_unknown_t = if_else(condition = is.na(source_unknown_t),
                                    true = 0,
                                    false = source_unknown_t)) %>% 
  mutate(adjust_live_weight_t = source_known_prop*source_unknown_t)
  #select(-source_known_prop, -source_unknown_t, -unknown_type)

check <- source_unknown %>% 
  group_by(consumer_iso3c, source_unknown_t) %>% 
  summarise(adjust_live_weight_t = sum(adjust_live_weight_t)) %>% 
  mutate(diff = source_unknown_t - adjust_live_weight_t)

# The biggest difference between the adjusted tonnes and unknowns is e-11 so it is negligible and just due to R rounding


```

### Combine proportioned unknown tonnes with the consumption_live_weight_t
```{r}
# Join all together
fmfo_matrix_new <- source_unknown %>% 
  # add consumption_live_weight_t and adjust_live_weight_t
  mutate(consumption_live_weight_t = consumption_live_weight_t+adjust_live_weight_t) %>% 
  select(consumer_iso3c, source_country_iso3c, dom_source, consumption_live_weight_t)

```

### Check consumption
Check to make sure the total consumption is not different for countries.
```{r}
# take average consumption over years data frame
fmfo_check <- fmfo_matrix_mean %>%
  group_by(consumer_iso3c) %>%
  # total consumption weight
  summarise(consumption_original = sum(consumption_live_weight_t, na.rm = TRUE)) %>%
  # join with adjusted data
  left_join(fmfo_matrix_new %>%
              group_by(consumer_iso3c) %>%
              # total consumption weight
              summarise(consumption_new = sum(consumption_live_weight_t, na.rm = TRUE)),
            by = "consumer_iso3c") %>%
  mutate(diff = consumption_new-consumption_original) %>%
  mutate(perc_diff = 100*(diff/consumption_original)) %>%
  arrange(desc(abs(perc_diff)))

```

Any differences in tonnage between the the original data and new data is extremely small (e+-14), this is likely due to rounding of precise numbers in R so nothing to worry about.

###  Save
```{r}

write_csv(fmfo_matrix_new, paste0(save_path, "fmfo_trade_matrix_w_unknowns_allocated.csv"))

```

