---
title: "Pressures Efficiencies workflow"
output: html_document
date: "2023-06-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal of script:
The pressures data reports the pressure efficiency based on the product group name (ex. cows_meat, yams_produce, soyb_produce, etc.). The trade data reports products in a variety of ways. The goal of this script is to pull in the different product names of from the trade data and pressures data to make a key for the items to be matched appropriately. Inland (freshwater) aquaculture is handled in its own script because the pressures efficiency data are global averages rather than specific to each country. See the script `02b_fw_aqua_pressures.Rmd`.

## Set Up
Packages
```{r}
library(tidyverse)

```


File Paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
pressures_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/pressures/"
matrix_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/01b_calc_trade_matrix_2017/"
seafood_trade_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/trade/aquatic_food_trade/"
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/02a_efficiencies_data_prep/"
```

## Read in the pressures data
```{r}
# read in the pressures efficiency data
pressure_efficiency_raw <- read.csv(file.path(pressures_path, "pressures_efficiencies.csv")) %>% 
  rename(nceas_group = category1_production)
```

## Read in and clean trade data to make matching keys
```{r}
#trade_data <- read.csv(file.path(trade_path, "trade_matrix_w_pressures_accounting_for_feed.csv"))

# read in the trade data calculated in 01b
trade_matrix <- read_csv(paste0(matrix_path, "trade_matrix_by_country_of_origin_2015_2019_import.csv"))

# read in data to map item_code from trade_matrix to pressures data
map_to_fao <- read_csv(paste0("/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/Codes/", "MapSPAM_to_FAO_v2.csv")) %>% 
  select(SPAM_super, SPAM_full_name, FAO_item_code, product_description_FAOSTAT)

# read in the fish trade categories to match with appropriate pressures
seafood_key <- read.csv(file.path(seafood_trade_path, "20230626_nceas_consumption.csv")) %>% 
  # retrieve the distinct combinations of habitat, method, and nceas_group
  distinct(habitat, method, nceas_group) %>% 
  # remove unknown nceas_group as we took care of these in 01c
  filter(nceas_group != "unknown",
         # we only want marine categories from the trade
         habitat != "inland") %>% 
  # assign organism based on nceas_group for seafood
  mutate(organism = case_when(nceas_group == "shrimps_prawns" ~ "shrimp",
                              nceas_group == "benthic" ~ "benthic",
                              nceas_group == "demersal" ~ "demersal",
                              nceas_group == "large pelagics" ~ "large-pelagic",
                              nceas_group == "medium pelagics" ~ "medium-pelagic",
                              nceas_group == "small pelagics" ~ "small-pelagic",
                              nceas_group == "reef associated" ~ "reef",
                              nceas_group == "salmonids" ~ "salmon",
                              nceas_group == "marine_fish_general" ~ "marine-fish-general",
                              nceas_group == "bivalves" ~ "bivalve",
                              nceas_group == "tuna" ~ "tuna",
                              nceas_group == "crustaceans" ~ "crustaceans",
                              nceas_group == "fofm" ~ "fofm")) %>% 
  # add in row for freshwater capture fish
  bind_rows(data.frame(habitat = "freshwater", method = "capture", nceas_group = "inland capture", organism = "fish")) %>% 
  # remove jellfish nceas group as it is not captured in trade data
  filter(!is.na(organism))
  
# read in the fao production item data to make livestock categories
livestock_FAO_categories <- read_csv(paste0("/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/Codes/", "FAOSTAT_production_item_codes_8-5-2022.csv")) %>% 
  # filter for meat products using the item code
  filter(`Item Code` > 865,
         # only need items that we have trade data for so use the item_code column from the trade_matrix
         `Item Code` %in% unique(trade_matrix$Item_Code)) %>% 
  # select the item code and item columns only
  select(item_code = `Item Code`, item = Item) %>% 
  # make the item strings all lower case
  mutate(item = tolower(item)) %>% 
  # assign nceas_groups to the correstponding item so that we are able to match livestock and their pressures
  mutate(nceas_group = case_when(str_detect(item, pattern = "eggs") ~ "chickens_eggs",
                                 str_detect(item, pattern = "cow") & str_detect(item, pattern = "milk") ~ "cows_milk",
                                 str_detect(item, pattern = "sheep") & str_detect(item, pattern = "milk") ~ "sheep_milk",
                                 str_detect(item, pattern = "goat") & str_detect(item, pattern = "milk") ~ "goats_milk",
                                 str_detect(item, pattern = "buffalo") & str_detect(item, pattern = "milk") ~ "buffaloes_milk",
                                 str_detect(item, pattern = "cattle") & str_detect(item, pattern = "meat") ~ "cows_meat",
                                 str_detect(item, pattern = "sheep") & str_detect(item, pattern = "meat") ~ "sheep_meat",
                                 str_detect(item, pattern = "goat") & str_detect(item, pattern = "meat") ~ "goats_meat",
                                 str_detect(item, pattern = "pig") & str_detect(item, pattern = "meat") ~ "pigs_meat",
                                 str_detect(item, pattern = "chicken") & str_detect(item, pattern = "meat") ~ "chickens_meat")) 
  


```

## Connect Pressure per Tonne data to the appropriate key

### Crops
We are adding the fao data key to the pressure efficiency data so that we will be able to connect the trade with the pressures. For crops, some of the SPAM_super items have multiple FAO item codes (ex. pota is the SPAM_super for potatoes and potato starch). These will therefor have the same associated pressure because of how the pressures data is aggregated. So when the final output has more rows than the original pressure_efficiency data it is because it is using the same pressure efficiencies for items under the same SPAM_super category.
```{r}
crop_pressure_per_tonne <- pressure_efficiency_raw %>% 
  # filter pressures data to produce
  filter(product == "produce") %>% 
  # join with map_to_fao key using organism and SPAM_super categories
  left_join(map_to_fao, by = c("organism" = "SPAM_super")) %>%
  # keep certain columns only
  select(iso3c, nceas_group, 
         # make FAO_item_code just item_code
         # rename to just item_code for matching with trade
         item_code = FAO_item_code, 
         # shorten to item
         item = product_description_FAOSTAT, 
         pressure, pressure_per_tonne) %>% 
  # add a food_group column for crops
  mutate(food_group = "crops")
```

### Livestock 
We are adding the fao data key to the pressure efficiency data so that we will be able to connect the trade with the pressures. For livestock items, some of the nceas_groups have multiple items associated with them (ex. chicken_eggs is for both "eggs, hen, in shell" and "eggs, other bird, in shell"). These will therefor have the same associated pressure because of how the pressures data is aggregated. So when the final output has more rows than the original pressure_efficiency data it is because it is using the same pressure efficiency for items under the same nceas_group category.
```{r}
livestock_pressure_per_tonne <- pressure_efficiency_raw %>% 
  # filter pressures data to meat (non-seafood) products from the organism column, buffaloes included to capture milk
  filter(organism %in% c("chickens", "cows", "goats", "sheep", "pigs", "buffaloes")) %>% 
  # join with livestock matching key
  left_join(livestock_FAO_categories, by = "nceas_group") %>% 
  # match columns with crop_pressure_per_tonne
  select(iso3c, nceas_group, item_code, item, pressure, pressure_per_tonne) %>% 
  # add food_group column to easily parse later
  mutate(food_group = "livestock")

na_test <- livestock_pressure_per_tonne %>% 
  filter(is.na(item))

unique(na_test$nceas_group)

```
We have pressures data for goats milk however we have no trade data that matches this. We will determine later if this is a problem when we join the two together

### Fisheries and Mariculture Pressure per Tonne
```{r}
fisheries_pressure_per_tonne <- pressure_efficiency_raw %>% 
  # filter for marine fisheries organism
  filter(organism %in% c("benthic", "demersal", "fofm", "large-pelagic", 
                         "medium-pelagic", "small-pelagic", "reef", "fish", 
                         "salmon", "marine-fish-general", "bivalve", 
                         "shrimp", "tuna", "crustaceans")) %>% 
  # remove nceas_group from pressures data as seafood key has appropriate nceas_group
  select(-nceas_group) %>% 
  #join with key
  left_join(seafood_key, by = "organism") %>% 
  # match with livestock and crop pressure per tonne dfs besides item and item code
  select(iso3c, nceas_group, pressure, pressure_per_tonne) %>% 
  # add in grouping
  mutate(food_group = "fisheries and mariculture")
```

### Bind pressures per tonne data
The fisheries and mariculture data will not have an `item` or `item_code` in `pressures_per_tonne` because they are matched with the trade data based on their assigned `nceas_group`. Some crop data may be missing an `item` description because the pressure data or trade data did not have one.
```{r}
pressures_per_tonne <- crop_pressure_per_tonne %>% 
  bind_rows(livestock_pressure_per_tonne) %>% 
  bind_rows(fisheries_pressure_per_tonne)

write.csv(pressures_per_tonne, file = paste0(save_path, "pressures_per_tonne.csv"), row.names = FALSE)
```


