---
title: "02ai_fw_aquaculture_pressures"
output: html_document
date: "2022-11-02"
editor_options: 
  chunk_output_type: console
---

GOAL: Calculate pressures for freshwater aquaculture by country and taxa group.
```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(countrycode)
```

File Paths
```{r}
raw_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/"
pressures_data_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/raw_data/pressures/"
save_path <- "/home/shares/food-systems/social_justice_projects/food_trade_pressures/output_data/02b_fw_aquaculture_pressures/"
```

# Join with on-farm stressors
Goal is to have stressors into per tonne units that match the other pressures data. The units of the freshwater aquaculture pressures data is the following:

- kg N/P per tonne (nutrient)
- m3 water per tonne (water)
- kgC02eq per tonne (GHG)
- m2 per tonne (disturbance)

We need to convert the pressures to match the units of the other pressure efficiency data:

- tonnes N/P per tonne (nutrient) - 1000 kg/ tonne
- m3 water per tonne (water) - no change
- tonnes C02eq per tonne (GHG) - 1000 kg / tonne
- km2 per tonne (disturbance) - 1000000 m2 / km2

For now, we are excluding the milkfish pressure values as they are quite different compared to other misc diadromous fishes.

```{r}
# read in environmental pressures data for freshwater aquaculture
pressures <- read.csv(paste0(pressures_data_path, "fresh_water_aqua_pressures_Jessica/SI_stressor_results_on_off_farm.csv")) %>%
  filter(weight=="live") %>%
  filter(production=="aquaculture") %>%
  filter(source == "on-farm") %>%
  filter(allocation == "mass") %>%
  # for now remove milkfish
  filter(taxa != "milkfish") %>% 
  # select stressor_group to match with tonnes_lw data, stressor, and median
  dplyr::select(stressor_group = taxa, stressor, median)

# make df for conversions of stressor data
conversions <- data.frame(stressor = c("GHG", "N", "P", "Water", "Land"),
                          # convert using the units as outlined in the description above
                          conversion = c(0.001, 0.001, 0.001, 1, 0.000001),
                          pressure = c("GHG", "nutrients", "nutrients", "water", "disturbance"))

# join pressures data with conversions
pressures_converted <- left_join(pressures, conversions, by = "stressor") %>%
  # convert pressures data to per_tonne using conversions
  mutate(pressure_per_tonne = median*conversion) %>%
  # select the stressor_group (taxa), pressure type, and pressure per tonne
  select(stressor_group, pressure, pressure_per_tonne) %>%
  # calculate the pressure_per_tonne total for each stressor group and pressure
  group_by(stressor_group, pressure) %>%
  summarize(pressure_per_tonne = sum(pressure_per_tonne)) %>%
  ungroup()

# Save data
write_csv(pressures_converted, paste0(save_path, "fw_aquaculture_pressures_per_tonne.csv"))
```

